### 2016-05-16 classes/mse330/final/final.tex

% !TEX program = xelatex
\input{../../../starter.tex}
\setmainfont[
Ligatures=TeX,
Extension=.otf,
UprightFont= *-regular,
BoldFont=*-bold,
ItalicFont=*-italic,
BoldItalicFont=*-bolditalic]{texgyretermes}

\title{The Effect of Bail on Guilty Pleas\\
\large{\textit{MS\&E 330}}}
\author{
Ali Alkhatib, Nikhil Garg, Martin Lindsey, Danaë Metaxa--Kakavouli
\\
\{\href{mailto:al2@stanford.edu}{al2},
  \href{mailto:nkgarg@stanford.edu}{nkgarg},
  \href{mailto:mlindsey@stanford.edu}{mlindsey},
  \href{mailto:metaxa@stanford.edu}{metaxa}\}@stanford.edu}

\bibliographystyle{apalike}

\begin{document}

\begin{titlingpage}
  \maketitle
\end{titlingpage}

\begin{abstract}
Research on the subject of
bail,
the burden it places on defendants, and
the persuasive and perhaps coercive influence it bears on the justice system
has been explored to some extent ethnographically, but thus far
researchers have not had the data necessary to conduct rigorous quantitative studies on
bail--setting and
the factors which influence pleading behavior. % and the influencing factors therein have not had sufficient data
With a unique insight into the demographics and characteristics of defendants and their outcomes,
we explore two questions:
\begin{inlinelist}
\item Does a greater burden, as measured by bail, cause people to plead guilty? And
\item Are judges handling cases uniquely, or uniformly?
\end{inlinelist}
Our findings reveal that heightened burdens do seem to cause defendants to plead guilty at a higher rate,
and that there is some evidence that variance in bail--setting is unique to judges.
\end{abstract}

\section{Introduction}
\input{intro}

\section{Literature Review}
\input{related_literature}

\section{Method}
\input{method}

\section{Results}
\input{results}

\section{Discussion}
\input{discussion}

\section{Conclusion}
\input{conclusion}

\input{appendix}

\newpage
\bibliography{../../../references.bib}

\end{document}