### 2016-05-26 cscw/papers/2017/labor/trackingWork.tex
subsection 2 is more or less fleshed out; ss3 is still not really done

\documentclass{sigchi}

\usepackage{todonotes,txfonts,balance,graphics,color,alphalph}
\usepackage{booktabs,textcomp,microtype,ccicons}
% \usepackage{babel}
% \usepackage{csquotes}
\usepackage[citestyle=numeric,backend=bibtex,bibencoding=ascii]{biblatex}
\usepackage{enumitem}
\usepackage{moreenum}
\usepackage[T1]{fontenc}
\usepackage{nth}
\usepackage[pdftex,pdfpagelabels=false]{hyperref}
\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage[utf8]{inputenc} % for a UTF8 editor only

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
  label=\arabic*),
}

\def\plaintitle{Naming Things is Hard: Real Title Following Colon}
\def\plainauthor{All the people (Ali, Margaret, MSB, who else?)}
\def\emptyauthor{}
\def\plainkeywords{Please don't make me pick keywords.
This is like asking a teacher to give the bullet points of
what a student missed in lecture.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}
\definecolor{PineGreen}{HTML}{008800}
\definecolor{BrickRed}{HTML}{FF0000}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{BrickRed}[\textit{al2: #1}]}}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}
\bibliography{../../../../references}

% End of preamble. Here it comes the document.
\begin{document}
\title{\plaintitle}

\numberofauthors{3}
\author{%
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
  \alignauthor{Leave Authors Anonymous\\
    \affaddr{for Submission}\\
    \affaddr{City, Country}\\
    \email{e-mail address}}\\
}

\maketitle

\begin{abstract}
  With growing attention on gig work
  --- ranging from the ``sharing economy'' to microtasks ---
  scholars have made connections to frameworks like Taylorism,
  and mechanisms such as worker advocacy and empowerment,
  to make sense of our observations of
  on--demand work and the workers that power this movement.
  We argue that our the underlying trend towards ``piecework''
  --- driven in part by the discretization,
  routinization,
  and external management of said work ---
  not only suggests, but in fact generates what we have observed:
  members of this transient workforce increasingly feeling
  disempowered,
  marginalized,
  and frustrated by the systems and platforms on which they work.

  After evaluating this framing
  through a series of case studies
  in various industries falling broadly under the ``gig work'' category,
  we turn our theoretical lens to look to the future,
  to identify worthwhile questions and
  points of inquiry that researchers in social computing should consider
  as we attempt to anticipate and perhaps shape the future of work.
\end{abstract}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section{Introduction}
The past decade has seen
microtasks, the ``sharing economy'', and other instantiations of on--demand contract work
grow to occupy the fascination of both academic circles and our culture as a whole
\cite{Ross,CrowdsourcingUserStudies,scholz2012digital}.
The research community has made connections between this emergent form of work
and the historically situated ``piecework'',
as well as a number of observations relating to
the frustration workers feel stemming from the management of this work
\cite{turkopticon,dynamo,uberAlgorithm}.
But the connections between
historical piecework and
contemporary crowd work
haven't been deeply made as yet.
% um, note to self: say more.

While much of this work appears to gesture toward
the parallels between contemporary on--demand work and piecework,
it's proven difficult to bring the totality of these observations into focus using one theoretical lens.
This paper will attempt to do so by arguing that
the topics social computing researchers have investigated are
not just parallel to historical piecework and
the process of factorization that took place in the early \nth{20} century;
indeed, these phenomena precipitate and reinforce each other.

In the reflection on the literature published in the last 5 years since
\citeauthor{crowdworkFuture}'s 
\citetitle{crowdworkFuture},
we notice a broader trend describing the change in work that's being done
\cite{crowdworkFuture}.
Following the improvements in
breaking down tasks,
delegating work, and
managing workers more broadly,
Crowd--powered has continued to parallel historical piecework's trajectory
by outgrowing ad--hoc worker groups and coalescing into more formal working groups;
we might call this the beginning of factorization.
% "Factorization" is not the right word but I honestly have no idea what the right term is.
% The process of formalizing work into factories? Should I tweet/post/ask about this somewhere?

% Crowd work,
% they argued in \citeyear{crowdworkFuture},
% represents a threat to the future of work inasmuch as it marginalized
% and perhaps even harmed
% workers;
% their driving question, then, was
% whether the future of
% (crowd, but arguably all discretized)
% work might be one in which they wish their children to participate someday;
% it seems strikingly like one that might have been asked
% as patches of denim were first being delivered to the homes of early pieceworkers.
% agreed; i'm commenting it out for now to see how I feel without making this argument at all. If I really feel like it's missing, I should find evidence that contemporaries discussed it...

Making sense of this 
by looking at the movements toward
distributing work,
routinizing and breaking down tasks,
and externally managing workers
as linked to one another, 
we give ourselves a framing of contemporary piecework that explains,
and arguably predicts,
what we have seen thus far
--- and perhaps what we should expect to see going forward.

This paper will attempt to demonstrate that the familiarity is more than passing;
that the phenomena researchers have observed in on--demand digitally mediated labor markets
were inevitable milestones in the birth and life of piecework.
We will set out to show that this perspective of crowd work as an embodiment of piecework
predicts the myriad outcomes of contemporary crowd work, including
our developments of work--flows further abstracting work,
the troubling effects of those developments on factors such as pay and work quality, and
many if not all of the frustrations researchers have discovered among workers
in the study of crowd and gig work.

``But'',
as \citeauthor{scholz2012digital} points out,
``it would be wrong to conclude that
in the realm of digital labor there is nothing new under the sun''
\cite{scholz2012digital}.
For one, information work
(e.g. AMT \& Upwork)
and digitally managed work in general make it substantially easier
to keep workers geographically dispersed
compared to the factory workers that followed home--based piecework.
This aspect of work renders workers virtually invisible to the task solicitors
--- or ``requesters'' ---
despite being significantly more connected than historical pieceworkers were % this is a weak point...
\cite{turkopticon}.
Secondly, while
factory foremen and other middle--men
traditionally mediated the relationships between workers and managers,
today the visible agents are \textit{systems};
rather than employing individuals who ``personified the functions of management''
and can thus negotiate workers' needs,
socio--technical systems mediate these interactions % often with unrewarding outcomes
\cite{wray1949marginal}.
% \cite{storiesIraniSilberman}.
Finally, this characteristic of modern piecework makes
worker coordination for collective advocacy and action significantly more difficult,
necessitating special consideration to maintain the inertia of collective will
while focusing that energy productively
\cite{dynamo}.
\ali{I really don't like this paragraph\dots}
% here are undoubtedly more examples of differences.
% How do I not stake a claim that this is comprehensive?
% Or should I really flesh this out and make it (comprehensive)?

\subsection{Piecework as a lens to understand gig work}
\citeauthor{crowdworkFuture}
investigate the future of crowd work by
situating and interrogating it through ``piecework'',
which historically described work done in the home,
in manageable tasks,
often involving clear instructions
and payment only for work completed, not work done
(the differentiation, here, being that
one would be paid for the \textit{output} of the work,
not the \textit{duration})
\cite{crowdworkFuture}.
% \msb{is that true? by that definition, design is piecework too, right, because I get paid for the job, not for hours worked. But design doesn't seem like piecework.}.
% This work was largely in textiles and was predominantly done by women,
% which by itself is a topic worthy of unpacking
% \msb{instead of gesticulating, can we indicate what the punchline of unpacking it was?}
% \ali{let's just leave this in the graveyard for now; i feel like someone who knows what they're talking about with regard to feminist and gender studies
% (e.g. the Bardzells, Irani) would really do justice here, but i'd be getting in over my head}
% \cite{scott1975women}.
Given the scope,
we can frame piecework and gig work as sharing these important similarities:
\begin{inlinelist}
\item this form of work began in the home
\item the worker is paid for each discrete piece of work done, regardless of time or effort; and
\item the worker's status
(not only socially, but also economically)
is ambiguous, or at least the subject of some controversy.
% \msb{my first reaction was ``this isn't definitional''. But now I look back and see that you're just saying there are similarities. It would be helpful if the sentence made it clearer that the similarities were in not just what the work was but also in the environment around it.}
\end{inlinelist}.

\subsection{Context}
Importantly, since the reemergence of piecework we've seen substantial frustration
and resistance among the workers in this area
\cite{uberAlgorithm,turkopticon,dynamo}.
This paper attempts to make sense of the broader research on this piecework,
or ``gig work'', by framing this as one of several steps in the marginalization of workers,
% \msb{this feels out of order to me. I don't think resistance is a step in the marginalization of workers, isn't it a reaction to the marginalization of researchers?}
%
% this needs to be re-written to reflect the new split
%
%
% starting with the discretization of tasks,
% followed by routinization and the rise of work--flows,
% and finally the external management of workers.
% 
All of this is to say that these milestones follow sequentially,
not coincidentally but necessarily,
and that by tracing this path using the corpus of scholarship on labor and workers
we can both make sense of past events and perhaps reasonably predict next steps.
% \msb{I don't think we've offered much proof that this is the case yet. Why must they follow necessarily? Can you offer a cause and effect relationship?}

We explore each topic
--- discretization, routinization, \& management ---
by looking at case studies in social computing.
Having validated this lens as a way of reasoning about contemporary piecework, or ``gig work'',
we turn to look ahead, envisioning future areas that researchers in social computing
--- and particularly digitally mediated work ---
should explore.

\section{Case Studies}
The existing body of research has shed light on on--demand from various perspectives,
and revealed a number of topics that,
through our framing, are clearly situated together.
Those topics are, at a high level, as follows:
\begin{enumerate}
\item the processes involved in making work into tasks, or \textbf{discretization},
in ``\nameref{sec:MakingGigsWork}'';
\item the outcomes of that discretization, both on the work itself as well as the workers,
in ``\nameref{sec:Fallout}'';
and finally
\item the results of these forces; namely the turning of work toward \textbf{factories},
which we will discuss in ``\nameref{sec:Factorization}''.
\end{enumerate}

We will follow these topics, using them as prompts for case studies in
the emergence and
development of contemporary piecework.

\subsection{Making Gig(s) Work}\label{sec:MakingGigsWork}
% this is phase 1: the decomposition & abstraction of work;
% all of the research that has gone into work--flows and
% all the things we can do with crowd work.

This section will largely discuss the processes at work that make distributed,
digitally managed work both possible and indeed preferable for ``requesters''
(in other words, the employers who solicit workers).
This body of research spans a broad field
within the CSCW and broader HCI community.
In this context,
we'll look at the body of research through the lens of
highlighting contributions which expand what we
(as ``requesters'' of work) can do by managing workers in novel ways.
This work broadly consists of three areas:
\begin{inlinelist}
\item \nameref{sec:decomposition},
\item \nameref{sec:workAbstraction}, and
\item \nameref{sec:flexibility}
\end{inlinelist}.
As we explore this work, we'll
attempt to relate the advances in the design of crowd work to
the research contributions made in the research of assembly line manufacturing,
Taylorism, and
scientific management 
during the \nth{20} century
\cite{hu1961parallel}.

\subsubsection{Decomposition}\label{sec:decomposition}
Scholarship describing and exploring
the decomposition of tasks is perhaps the most established of the above areas among HCI researchers;
\citeauthor{bernsteinSoylent} made early contributions to this area with \textit{Soylent}
(by coining and employing the ``find--fix--verify'' approach)
\cite{bernsteinSoylent}.
\citeauthor{cheng2015break} found that microtasks
--- though not necessarily \textit{faster} than ``macrotasks'' ---
yield higher quality work,
particularly when that work might be readily interrupted
\cite{cheng2015break}.
\citeauthor{selfsourcingTeevan2014} further push the boundaries of decomposed work
by exploring ``selfsourcing'', and further this work with \citeauthor{selfsourcingTeevan2016}
\cite{selfsourcingTeevan2014,selfsourcingTeevan2016}.
While this work doesn't strictly fall under ``crowdsourcing'',
the major contributions here
seem uncontroversially to be inspired by the design of crowd work.

Much of the research in the space of designing crowd work has
sought to illustrate the potential to take highly creative or skilled work
and generate high--quality results.
Perhaps the most notable case study here can be found in
\citeauthor{foundry}'s \textit{Foundry}, which employed
``flash teams'' to achieve expert--level outcomes via thoughtful
decomposition of work as ``modular tasks''
\cite{foundry}.

Work decomposition is far from new;
``decomposition'' generally illustrates the same concepts of work that ``Taylorism''
and scientific management sought to embody
---
\citeauthor{silberman2010ethics} in particular foresaw this danger and warned of it in
\citeyear{silberman2010ethics}
\cite{crowdworkFuture,silberman2010ethics,nickerson2013crowd}.
In both the historical and contemporary cases of decomposed work,
work was,
at least initially,
distributed in the form of tasks to the homes of workers;
\citeauthor{riisOtherSideLives} captured this in his documentary work
\citetitle{riisOtherSideLives}
in \citeyear{riisOtherSideLives}
\cite{riisOtherSideLives}.

% In both the historical and contemporary cases,
% distributed, speculative work
% has promised
% to diffuse the costs of capital--intensive resources
% (whether they were sewing machines or laboratory equipment).

\subsubsection{Work Abstraction}\label{sec:workAbstraction}
Decomposition allows requesters to assign tasks without concern for the broader context.
While we'll discuss this aspect of crowd work more critically later,
it's worth pointing out that discrete blocks of work containing all the relevant context for a worker
allows workers to engage with virtually any component of work without worrying that their lack of 
higher--level awareness of the goals of the requester might negatively affect their work.

\citeauthor{chilton2013cascade} perhaps best illustrated this with
\textit{Cascade} by demonstrating that it's possible to
break certain classes of tasks apart
in such a way that they yield taxonomies of various subjects,
a task previously thought to be safely within the domain of expert workers
with top--down awareness of the context of the work as a whole
\cite{chilton2013cascade}.
\citeauthor{verroios2014context} further illustrate this potential by
forming a task one might consider highly contextually dependent
--- summarizing the contents of a movie ---
in such a way that crowd workers could contribute small pieces of work without
needing to know the content of the rest of the project
\cite{verroios2014context}.

Here, \citeauthor{hu1961parallel}'s work,
saying of assembly line work that
``It is assumed that men are of equal ability and every man can do any of the $n$ jobs'',
parallels the approach that dominated early research into crowd work
--- namely, using non--expert crowds for complex work
\cite{hu1961parallel}.
This mindset in \citeauthor{hu1961parallel}'s analysis,
and indeed the study of factory and mass manufacturing labor through the \nth{20} century,
substantively owes its existence to scientific management
and the rigorous decomposition of work into tasks, discussed earlier,
and persists to this day as it colors
researchers' goals and objectives in the study and design of crowd work.

Piecework's influence on the abstraction of work into tasks,
described above, is more than just caused by the decomposition of work;
work abstraction itself makes it possible for workers to come and go flexibly,
prompting work requesters to consider ways to design these now discrete tasks in ways that
maximize flexibility, both by allowing (and even anticipating) some inconsistency in worker availability
\textit{and} allowing and anticipating some inconsistency in the quality of the work output itself.
It's to this area that we now turn our attention.

\subsubsection{Flexibility}\label{sec:flexibility}
Earlier we discussed \citeauthor{cheng2015break}'s work
measuring the impact that interruption has on worker performance.
This work both points to and embodies a broader sentiment in
both the study and practice
of crowd work that microtasks should be designed flexibly,
fully exploiting the abstracted nature of each piece of work
\cite{interruptionIqbal,delayAndOrderLasecki,vaish2014low}.
While \citeauthor{cheng2015break} identified costs with breaking tasks into smaller components
in the form of higher cumulative time to complete
(albeit much shorter real time to complete, owing to parallelization),
\citeauthor{delayAndOrderLasecki} found that at least \textit{some} performance can be recouped by stringing 
similar tasks together.

Given the importance of consistent work results, one might intuit that
requesters would prefer high--quality workers who can be relied upon to be available
(even for contextually independent tasks),
which would appear to contradict the benefits of flexibility already discussed;
requesters have thus made significant headway toward
``embracing error'' to allow requesters to maximize the benefits of a flexible,
even transient,
workforce.

\citeauthor{embracingErrorKrishna} offer orders--of--magnitude improvements
in various binary classification tasks
% --- a typical application crowd work ---
on the principle that myriad diverse workers complete these tasks
in order to accurately inform the model.
And rather than building tasks to \textit{tolerate} worker drop--off and attrition,
some researchers have designed work predicated on the expectation of this phenomenon:
\citeauthor{sensitiveTasks} describe ways of assigning tasks in such a way that
crowd workers would never be given enough information to piece together sensitive information about
any single topic
\cite{sensitiveTasks}.

\subsection{The Fallout of Crowd Work}\label{sec:Fallout}
% If the research exploring and documenting how we can exploit crowd \textit{work} can be described as wide--ranging,
% the scholarship discussing the ways crowd \textit{workers} have been exploited is more focused;
\citeauthor{turkopticon} point out the disillusion that companies such as Amazon foster on platforms for work like AMT
(see also \citeauthor{dynamo}'s work
continuing in the spirit of this observation to generate collective action to improve worker conditions)
\cite{turkopticon,dynamo}.
\citeauthor{uberAlgorithm}
find similarly that workers on gig work platforms are frustrated by the systems on which they work,
to say little of the policies which these systems enforce
\cite{uberAlgorithm}.

We discussed the benefits of flexibility
(both in the sense of having arbitrary workers perform tasks and
in the sense that we can design tasks to be more resilient to poor work)
in the previous section.
It's from that point in the literature that we turn our attention to
the perhaps unintended effects of crowd work
and the affordances for transience that we build into this mode of work.
We'll address two major areas of work under this subject:
\begin{inlinelist}
\item \nameref{sec:lowPay}; and
\item \nameref{sec:varQualWork}.
\end{inlinelist}

\subsubsection{Low Pay}\label{sec:lowPay}
\citeauthor{laborEconomicsOfCrowdsourcingHorton}
identified problems with crowd work wages relatively early on,
attempting to address this imbalance from a behavioral economic perspective ---
that is, identifying and presenting a model that describes a worker's
``\textit{reservation wage}''
\cite{laborEconomicsOfCrowdsourcingHorton}.
This work has largely informed much of the research into and practice of estimating crowd work compensation
\cite{incentivesShaw,paolacci2010running}.
More cynically,
we would describe this work as identifying optimal levels to activate race conditions among piece workers.

But we turn to \citeauthor{turkopticon}'s discussion of ``\textit{Turkopticon}'',
a system they designed to interrogate worker invisibility and to promote better wages across several dimensions
\cite{turkopticon}.
Of particular relevance here,
\citeauthor{turkopticon} call to attention that ``Turkers'' are ultimately vulnerable to
wage theft and
pay rates that translate to well under minimum wage.
Returning to \citeauthor{laborEconomicsOfCrowdsourcingHorton},
we find that the median ``reservation wage'' in \citeyear{laborEconomicsOfCrowdsourcingHorton}
was \$1.38, while the mean was \$3.63
\cite{laborEconomicsOfCrowdsourcingHorton}.

Understanding workers' motivations given these conditions has thus become a goal for some researchers
\cite{whyWouldAnyoneBrewer}.
\citeauthor{Sun20111033} conclude that
``\dots~solvers participate in online tasks
not only for money
but also for enjoyment
or the sense of self--worth''
\cite{Sun20111033}.
This might have rung true in \citeyear{Sun20111033},
and certainly corroborates \citeauthor{Ross}'s findings after investigating
``who are the crowdworkers'',
but as \citeauthor{whoareNOTtheTurkers} points out
``we [have since] learned that most tasks on AMT are done by a small group of professional Turkers\dots''
\cite{Ross,whoareNOTtheTurkers}.

Now, \citeauthor{turkopticon}
and later
\citeauthor{dynamo} cite insufficient pay as a central point of frustration among workers,
via \citeauthor{irani2015cultural} and \citeauthor{dawnDigitalSweatshopCushing}'s work
\cite{dynamo,irani2015cultural,dawnDigitalSweatshopCushing,turkopticon}.

\subsubsection{Variable quality work}\label{sec:varQualWork}
Meanwhile,
researchers have struggled with what we might call ``variable quality work''
along two dimensions:
First, to use the characterization of one of these contributions,
``understanding malicious behavior''
\cite{MaliciousCrowdworkersGadiraju}.
While some work has cast workers as ``malicious'' agents,
the evidence thus far suggests that
workers behave in unexpected ways as they attempt to assert some control over their interaction with the system
(a topic of discussion to which we'll return later)
\cite{uberAlgorithm}.
The second dimension of research in this space generally attempts
to eke out the highest quality work possible from workers
given the apparent difficulty in predicting work outcomes

% Low pay yields variable quality work for a number of reasons,
% but before we discuss the causes of this effect, we should discuss the

% Variance in the quali

The effect that poor wages have had on piece work and factory workers is hardly unknown or novel;
\citeauthor{gantt1913work} discuss this exact mechanism in his book on
\citetitle{gantt1913work}, pointing out that
``\dots~where there is no union,
the class wage is practically gauged by the wages the poor workman will accept,
and the good workman soon becomes discouraged and \textit{sets his pace by that of his less efficient neighbor},
with the result that the general tone of the shop is lowered'' (emphasis added)
\cite{gantt1913work}.
\ali{Do I need to follow up or does this basically speak for itself?}

This research is similar to, but subtly different from, the notion of the ``market for `lemons'''
which \citeauthor{fort2011amazon} discuss;
specifically, \citeauthor{akerlof1970market}'s writing of a ``market for `lemons'''
describes a marketplace where the quality of the product or service is unknown to the buyer
\cite{fort2011amazon,akerlof1970market}.
The effect of this \textit{perceived} uncertainty is that
the \textit{actual} trustworthiness drops precipitously
as all of the consistent, reliable, high--quality workers capable of leaving these markets do so,
leaving only the ones who cannot or will not establish their trustworthiness.

Suffice it to say, then, that poor pay and poor work are linked,
and that we should not be surprised to find this relationship play out online as strongly as it does offline.
Indeed,

% \itshape
% sections here should include:
% \begin{enumerate}
% \item ``flexibility'' \cite{friedman2014workers,DesigningForLabourUber}
% \item leads to ``low pay'' \cite{turkopticon,dynamo}
% \item leads to ``variable quality work'' \cite{MaliciousCrowdworkersGadiraju,AccuracyTradeoffsLasecki}.
% \end{enumerate}
% \upshape

\subsection{Factories}\label{sec:Factorization}

\rule{\linewidth}{1pt}

% \textit{There's a reference to something that's \textbf{not} 99designs,
% that describes a call for scientific solutions/designs to solve some problem
% that the requesters have.
% Honestly it's killing me that I can't remember the name,
% and I need to find this before I do anything productive ever again.}
% \msb{netflix prize?}
% that's an interesting example in the same vein as what i'm thinking about, but this is a whole business of it
% \textbf{INNOCENTIVE} oh my gosh it just clicked my life is complete right now at 9:54am.
% I don't even need to submit anymore.
% I'm satisfied having pulled this strand of a memory out of my brain.

The HCI community is perhaps most familiar with examples such as
Innocentive,
Uber, Lyft,
and Amazon Mechanical Turk (AMT),
which all allow requesters in various forms to tap into
resources such as
cars,
computers, and above all
``cognitive surplus''
with relative ease
\cite{howe2006rise,DillahuntPromise,storiesIraniSilberman,shirky2010cognitive}.
This insight, that
workers can be geographically distributed and tasks decomposed,
% (especially given careful central management of those workers, discussed later),
has proven remarkably compelling
and an effective fulcrum for leveraging the Internet for highly scalable work
\cite{sensitiveTasks,embracingErrorKrishna}.

% The insight here seems to have been that
% a substantial amount of work does not necessarily need to be done in collocated space,
% but can be done from one's own home or other space arranged by the worker.
% Moreover, by shifting the workplace into the worker's domain,
% the cost of capital involved in that work can more easily be placed on workers' backs
% \msb{I don't think that was the original motivation for AMT or Uber --- it was about reaching people who could not otherwise join the labor force for Amazon's data cleaning or for taxi-ing.}.
% I didn't mean that this was some insight that Uber or Lyft had, but that this was some part of what makes Uber work;
% i would argue categorically that these companies wouldn't functionally work if they weren't able to obviate the 
% maintenance costs that a factory owner would associate with their capital (e.g. auto costs)
While yellow--cab organizations often manage drivers in similar ways to platforms such as Uber and Lyft
(inasmuch as drivers are often considered contractors, or otherwise exempt from the formal status of ``employee''),
these new systems go further
by framing ownership and responsibility for costs
(e.g. automotive maintenance)
on drivers.
In our fieldwork, we discovered many longtime ``on--demand'' drivers
(affiliated with platforms such as Uber and Lyft)
who readily voiced frustration with the costs of
frequent oil changes and the unexpected replacement of expensive parts.
Distributing labor had similar effects on pieceworkers at the turn of the \nth{20} century;
textile workers were given the source material with which to work,
but were generally expected to use their own sewing materials and work in their own homes
\cite{hapke2004sweatshop}.

% \subsection{Discretization and Routinization}
% \subsubsection{Decomposition}\label{sec:decomposition}
Distributing work in method described above may be thought of as vertically slicing work such that 
each person is responsible for the whole task
--- making a whole garment or being responsible for driving in a single neighborhood.
Broken down in this way, work could grow to unprecedented scales,
but the quality of the work would remain relatively variable.
Textile work being a salient example,
some seamstresses might be better than others at the same task,
making this framing of the work initially problematic.

A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
manageable routines that could be taught relatively easily,
and whose work output could be evaluated in abstraction from the rest of the work.
In Ford's assembly line, this meant that workers were not responsible for building a whole car,
but a single very narrowly defined action that needed to be done on every car.

This approach paralleled what would be known as ``Taylorism'',
and its influences can be seen today in crowd work and micro--tasks.
\textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
horizontal cuts instead of vertical}
\cite{cheng2015break,writingMicroTasks}.
By decomposing and recomposing tasks,
and in particular by assigning similarly natured work to the same workers,
% and especially as the nature of the work could be aligned with other, similar tasks,
workers could become ``experts'' in a small aspect of the work that they did,
speeding their work dramatically
\cite{delayAndOrderLasecki}.
Perhaps more important, however, was that
the breaking down of work into tasks has made it more practical to evaluate work at each stage
[\textit{citation needed}].

% Taylorism enabled
% --- and would later become synonymous with --- 
% scientific management, which we will discuss later
% and might generally be regarded collectively as ``scientific management''
% The compelling solution that emerged in piecework became what we now know as Taylorism;
% rather than

% I never talk about factorization it's too soon ahhhh
% With piecework and factorization making mass manufacturing more viable,
% businesses could produce at scales that were previously unimaginable.
% Nevertheless, this approach to labor and the production capacity it offered was not without its hurdles.
% One of the issues that challenged mass production at the turn of the \nth{20} century,
% as well as today in crowd work and elsewhere,
% has been that of ensuring that the quality of the work is sufficiently good.

% hey so i need to bring in citations on piecework for this stuff because I didn't exactly live through this time

% is ANY of this good?

% The approach to piecework described above hinged in some sense on
% the shared understanding of the work to be done;
% workers generally understood what was required of sewing together pieces of fabric.
% Still, the work was speculative in that inadequately sewn clothing would not earn payment.
% This represented a risk for pieceworkers,
% who might spend time working on 10 garments only to be paid for 2 or 3.
% It also represented a potential cost for management,
% who would need to salvage the material
% (if it was even possible),
% and give that material to another pieceworker to do it correctly.

% i need to find the source for this, because it's honestly only a vague quote in my head,
% so... maybe i imagined it?

% One of the solutions that

% Broadly speaking, the solution to this problem is to make instructions clearer and easier to follow.
% Further, by establishing a shared understanding
% --- a ``collective conscious'' or even a corporate culture regarding work ---

% \textit{I want to argue that modern gig work is discretized,
% and that this parallels some of the piecework that we're historically familiar with}

Opening the door to more discrete modules of work known as tasks 
% Work has become discretized in the last 10 years, particularly in the information work sector.
Contributions such as
``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
and
``Cascade'' % (described in \citetitle{chilton2013cascade})
make it easier to assign smaller components of work to myriad workers and
recompose the constituent parts into something more complex
\cite{bernsteinSoylent,chilton2013cascade}.
Some work has taken special interest in the process of breaking tasks down into more manageable parts
% Some work has even turned to break it down'',
inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
both for the purposes of crowd--sourcing in the sense of out--sourcing work
as well as ``selfsourcing''
\cite{cheng2015break,selfsourcingTeevan2014}.

The parallels between this kind of work and
historical piecework of the turn of the \nth{20} century
is by no means new.

The \textbf{discretization} and decomposition of work certainly facilitates
breaking tasks down,
parallelizing work,
and getting more done across a broader array of people;
but the key advantage in turning macro tasks 
into many micro tasks is that we can make that work \textit{routine}.

By making larger bodies of work less contextually situated,
we can define processes that make that work
and the instructions thereof
useful for virtually anyone with shared cultural or intellectual background.

routinization can be seen in the contemporary in work--flows
\cite{foundry,bernsteinSoylent}.
% [\textit{There are various papers that talk about creating a work--flow;
% maybe this is where FFV belongs? Foundry?}].

% \textit{Harken back to piecework
% and the instructions that denim workers had to make clothing;
% the instructions were clear and scoped enough that you could repeat the task ad infinitum.}

% I commented everything going forward out not because it's all trash (it's not!)
% but because we pretty significantly reframed the cases to be studying and
% i didn't want the structure from before to influence the rewrite;
% we should be bringing content from below back above. just, judiciously...

% \subsection{Discretization and Routinization}

% Distributing work in method described above may be thought of as vertically slicing work such that 
% each person is responsible for the whole task
% --- making a whole garment or being responsible for driving in a single neighborhood.
% Broken down in this way, work could grow to unprecedented scales,
% but the quality of the work would remain relatively variable.
% Textile work being a salient example,
% some seamstresses might be better than others at the same task,
% making this framing of the work initially problematic.

% A compelling solution emerged in the early \nth{20} century to break tasks down into discrete,
% manageable routines that could be taught relatively easily,
% and whose work output could be evaluated in abstraction from the rest of the work.
% In Ford's assembly line, this meant that workers were not responsible for building a whole car,
% but a single very narrowly defined action that needed to be done on every car.

% This approach paralleled what would be known as ``Taylorism'',
% and its influences can be seen today in crowd work and micro--tasks.
% \textit{talk about breaking tasks down both in general (Cheng) and even for the self (Teevan);
% horizontal cuts instead of vertical}
% \cite{cheng2015break,writingMicroTasks}.
% By decomposing and recomposing tasks,
% and in particular by assigning similarly natured work to the same workers,
% workers could become ``experts'' in a small aspect of the work that they did,
% speeding their work dramatically
% \cite{delayAndOrderLasecki}.
% Perhaps more important, however, was that
% the breaking down of work into tasks has made it more practical to evaluate work at each stage
% [\textit{citation needed}].

% Opening the door to more discrete modules of work known as tasks 
% Contributions such as
% ``find--fix--verify'' % (described in \citetitle{bernsteinSoylent})
% and
% ``Cascade'' % (described in \citetitle{chilton2013cascade})
% make it easier to assign smaller components of work to myriad workers and
% recompose the constituent parts into something more complex
% \cite{bernsteinSoylent,chilton2013cascade}.
% Some work has taken special interest in the process of breaking tasks down into more manageable parts
% inasmuch as decomposing what we might call ``macro'' tasks into microtasks,
% both for the purposes of crowd--sourcing in the sense of out--sourcing work
% as well as ``selfsourcing''
% \cite{cheng2015break,selfsourcingTeevan2014}.

% The parallels between this kind of work and
% historical piecework of the turn of the \nth{20} century
% is by no means new
% [\textit{citation needed --- who said this first? most notably? Turner?}] \msb{not familiar with it, but google scholar suggests this was an early one to talk about it \url{http://scholarship.law.berkeley.edu/cgi/viewcontent.cgi?article=1433&context=bjell}}.
% \textit{Elaborate on gig work, piecework, etc\dots}

% The \textbf{discretization} and decomposition of work certainly facilitates
% breaking tasks down,
% parallelizing work,
% and getting more done across a broader array of people;
% but the key advantage in turning macro tasks 
% into many micro tasks is that we can make that work \textit{routine}.

% By making larger bodies of work less contextually situated,
% we can define processes that make that work
% and the instructions thereof
% useful for virtually anyone with shared cultural or intellectual background.

% routinization can be seen in the contemporary in work--flows
% \cite{foundry,bernsteinSoylent}
% [\textit{There are various papers that talk about creating a work--flow;
% maybe this is where FFV belongs? Foundry?}].

% \subsection{External Management}
% \textbf{Discretization}, enabling \textbf{routinization}, thus allows us to
% manage work more meaningfully.
% Routinization in particular made it possible
% (and indeed preferable) to assume that people
% ``\dots are of equal ability and \dots can do any of the $n$ jobs''
% \cite{hu1961parallel}.
% \citeauthor{storiesIraniSilberman}
% find similarly that
% this work is only made to seem new by framing the units of work as the output of ``task--completion APIs''
% \cite{storiesIraniSilberman}.

% Crucially,
% \citeauthor{hu1961parallel}'s case study assumed that workers were similarly trained and capable,
% which in \citeyear{hu1961parallel}
% meant that workers would at least have to be trained in person, if not instructed as well.
% Throughout the industrial revolution, workers flocked to centralized locales of production,
% and physically situated factories made it both
% difficult to ignore the general pattern of mistreatment and
% relatively easy to organize.

% \citeauthor{storiesIraniSilberman} describe is that the ``APIs'' in the latter case

% \textit{I want to cite Justin's work on crowdsourcing effort to highlight that
% we can evaluate workers for the work they do with increasingly finite amounts of time;
% also, Ranjay's work that embraces failure because
% we can essentially algorithmically make sense of failure (let's call it noise)
% and re--capture signal.}

% \subsection{Resistance}
% So it's not surprising, then, that workers are frustrated:
% Turkopticon, Dynamo, and many other pieces of work point this out
% \cite{turkopticon,dynamo,uberAlgorithm}
% [\textit{also want to cite Brian's work from CHI this year}].
% In popular culture, reporters have inquired about
% what ``the future of labor unions'' will look like
% [\textit{Imagine I had cited an online article and
% it wasn't totally inappropriate for a conference}].

% \textit{The parallels here would be difficult to ignore;
% Lily and others have pointed this out,
% but it's worth thinking about the broader trend of
% resisting management,
% driven by routinization,
% enabled by discretization.
% Do that here.}

\subsection{Looking forward}
If we agree that
discretization,
routinization,
management,
and even resistance
necessarily follow one another according to this theoretical lens,
then we have to use it to attempt to envision
what comes next.

\textit{I've been going back and forth regarding how to frame this paper;
the approach that tries to look at
things that are similar vs things that are different
doesn't seem to work, but
I've left the thoughts here because it's not all completely bad.}

\section{Things stay the same}
How is gig work the same as it's been historically?

I think this section would be compelling to draw parallels between
the narratives drivers gave about the flexibility, autonomy, etc\dots
and that which we might have seen among pieceworkers
(predominantly women, who benefited from being able to work from home).

\subsection{Flexibility}
Are there cultural differences between the people that did piecework and
the people that do gig work now?
I'm not sure there are significant differences that have affected the outcomes so far.

Many of the workers to whom piecework appealed were mothers, wives, etc\dots
who mostly stayed at home for various reasons
(certainly largely it was cultural
--- women weren't afforded equal access to labor opportunities,
making in--home job opportunities not only compelling, but also one of few available options).

Gig workers are in some senses similarly constrained:
workers on Amazon Mechanical Turk
--- those that use it as a primary source of income, at least ---
report being homebound for various reasons (e.g. medical, parenting, etc\dots).
Society and circumstance have made it difficult
or impossible
to join the contemporary, conventional workforce;
gig work re--opens that door.

But there are differences; during our research over the summer,
we spoke to drivers on Uber and Lyft, cleaners, and other gig workers.
Many of them told us about their home lives
--- about children, spouses, and other commitments ---
to which they wanted to dedicate more time.
One driver
(let's call him Ra\'{u}l)
told me about how he drove for Lyft
after working as an inventory manager at a hospital for more than nine years.

I asked him why he quit that job and forewent
the benefits,
predictability,
and career growth opportunity
that his old job offered.
He told me that when his daughter was born,
he was overcome with a desire to spend more time with her.
No longer satisfied with work
where he often left before his daughter would wake up and return after she fell asleep,
Ra\'{u}l decided to start driving for Lyft,
because he could drive in the evenings when his daughter was asleep.

Other drivers reported similar benefits;
gig work affords its workers flexibility that conventional careers don't allow.

\subsection{routinization of work}
This leads into the next section, but I want to bring up the process of
making work about mass--manufacturing, at least inasmuch as
the instructions are the same for everyone.

The Internet arguably has made it much easier to broadcast those instructions,
but it's had this deeper effect of enabling some amount of back--and--forth
between the worker and the (algorithmic) manager.

\subsection{Taylorism}
The routinization of work makes it possible to measure that process,
optimize it for certain characteristics, and ultimately
lead to Taylorism and scientific management.
This is not new;
researchers have studied and written about the slow creep of
algorithmic management and 
discretization \& routinization of work tasks.

\textit{We hope to take a step back from the context in which this work is often applied, and
look for its place in the larger trends and theories
to make sense of the trends of gig work at large.}

Industrialization and the automobile assembly line makes this famous,
but piecework functioned on the principle that everyone was making similar
or identical
garments and other products.

Now, we see Turkers being evaluated on the outcome of their work conforming to norms,
sometimes bootstrapped,
as in Ranjay's talk on 
``Embracing Error to Enable Rapid Crowdsourcing'',
but more conventionally in work flows like ``Find--Fix--Verify''.

\textit{This might be an opportunity to reflect on
how pieceworkers internalized the work they were doing,
responded to the stress of the uncertainty of potentially rejected work,
etc\dots but I'm not familiar with research in that space.}

\subsection{The emergence of decentralized workplaces}
The practice of in--home piecework was consumed by
the centralization of factories
(the effect of which we'll talk about later, since
we can talk about how this made unions more practical),
but for a time many pieceworkers
at the turn of the \nth{20} century
worked out of their homes.
Strikingly, many of the gig workers
\citeauthor{crowdworkFuture}
discussed in
\citeyear{crowdworkFuture},
and indeed many more continue to do what we might call ``information work''
--- that is, work that predominantly demands human computation ---
but increasingly we're seeing the movement toward transient work that largely requires embodied presence
\cite{uberAlgorithm} (and others).

\subsection{What was this about?}
I was going to check in quickly and take a photo of your whiteboard
(hoping that you'd kept our conversation notes around)
but your room was \textit{literally} full of people lol.

\section{Things are changing}
The medium on which this work is being done
--- and to an extent the medium used to manage workers ---
has dramatically changed things as well, however;
workers are distributed around the world,
working out of their cars in the cases of livery services
(notably, never returning to a base of operations)
across and between cities as well as nations,
or in their homes
(paralleling the trend of piecework even more closely).

Trying to understand how gig work has differed from piecework should at least
start with looking at the different characteristics of the work involved.
After that, we should think about how
the demography and culture of the people engaging in this kind of work
have changed versus that of the pieceworkers.

\subsection{Differences in the work itself}
Gig work has all of the above similarities with piecework,
but there are key differences.

Piecework emerged at the turn of the \nth{20} century,
right at the time that telecommunications began to boom in the United States. % too far a reach?
\textit{Did telecom enable remote management?}

Gig work in its contemporary formation is largely mediated by
ubiquitously accessible digital media
(the Internet, telephony, etc\dots)
and importantly has relied on this technology to facilitate the remote management of workers
\cite{uberAlgorithm}.

\textit{Has the work fundamentally changed,
or are we just being managed remotely in different ways?
I'm not entirely sure.}

How does this work differ from the experience of being ``on--call'' that is so familiar to retail employees?
How does this work differ from the work in which truckers, taxi drivers, and other
independent contractors have been participating for decades?
We argue that the substantive difference in these markets is the speed of the market itself,
motivated by the technology which mediates it.
Because workers can be sourced and dispatched virtually instantly,
businesses that engage in this kind of work
(e.g. Uber, Amazon Mechanical Turk, etc\dots)
have taken to removing other bottlenecks,
like vetting workers upfront

\section{More deliberate work}
% \subsection{Differences in the people that do the work}
% ???

\balance{}
\printbibliography

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: