### 2016-07-12 random/final.tex

% % !TEX program = xelatex
\documentclass[10pt]{article}
\usepackage{balance,graphics,setspace,parskip,times}
% \usepackage{fontspec}
% \usepackage{xltxtra}
% \usepackage{xunicode}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}

\usepackage[tiny,compact]{titlesec}

% \usepackage{xesearch}
% \newcounter{words}
% \newenvironment{counted}{%
%   \setcounter{words}{0}
%   \SearchList!{wordcount}{\stepcounter{words}}
%     {a?,b?,c?,d?,e?,f?,g?,h?,i?,j?,k?,l?,m?,
%     n?,o?,p?,q?,r?,s?,t?,u?,v?,w?,x?,y?,z?}
%   \UndoBoundary{'}
%   \SearchOrder{p;}}{%
%   \StopSearching}

\setlength{\parskip}{.4em}

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{
  label=\arabic*),
}
\def\labelitemi{\ }
\newcommand*\elide{\textup{[\,\dots]}}
\newcommand{\sectitle}[1]{\textbf{#1}\\}

% \author{
%   Ali Alkhatib\\
%   ali.alkhatib@cs.stanford.edu\\
%   \textit{Couldn't find an envelope; will coordinate to meet next quarter (if acceptable)}
% }

% \title{Final Paper\\
% \normalsize{
%   \textit{Digital media are an egalitarian force in American society}\\
%   }
% }

\begin{document}
The future of work may resemble a hunter--gatherer mode of work;
foraging for work throughout the day to sustain ourselves
rather than allowing us the stability of careers that we know today.

Alex wakes up in the morning to his alarm and
checks his phone for notifications of non--urgent tasks for which he's qualified
--- for instance, doing a second or third round of revisions for an essay for which he's already given feedback ---
while eating breakfast.
He flips from microtask to microtask over cereal and juice,
indicating whether the people in the pictures he's flipping through are having an argument or not.
He has a vague sense that he's providing training data to some artificial intelligence project, but the details escape him.
Or he never sought after them.
He doesn't care.

Occasionally Alex gets a prompt asking if he can set aside time later in the week to do some tasks
--- in this case, committing to training someone to do better in technical interviews by doing mock interviews over Skype ---
adding a few more blocks of appointments to his week.
His calendar is mostly empty, but
he tries to consolidate appointments as much as possible so he doesn't have to wake up at all hours of the day.
% After agreeing to some of those tasks, populating more of his calendar over the next several days,
% he packs up his laptop and phone and heads out.

Alex finishes his breakfast, wraps up his last task, and starts packing his messenger bag for the day.
As he starts to head out, he taps through an app interface to see if he can get a cleaner to come by his place today.

Michelle? No, she's never on time\dots~\\
Eric? No, seemed a little shady\dots~\\
\textbf{Lupe} is available, and she was reliable.

Alex pings Lupe to see if she can swing by today and do the laundry that's been piling up in the hamper.

Alex's phone steadily buzzes every few seconds with notifications that people are looking for workers.
Some are looking for a ride to the coffee shop he's already heading to;
others are looking for a ride downtown;
still others are looking for a ride to the airport.
% Knowing that his intent is to go to Philz Coffee every morning and seeing that Alex has recently woken up,
% Uber starts suggesting to him people in his building or neighborhood who are interested in going in the same direction.
Today Alex wants to get some work and later meet with some friends, so he declines them all.
He doesn't like driving people around anyway;
passengers are rude and the pay isn't very good.

He makes his way to Philz ignoring the now streaming solicitations on his phone:
Pick up someone's medicine;
tutor someone's kid in math;
babysit someone's kid for a few hours on an abrupt need
(for \textit{4 times} the normal value because of the urgency of the task request);
eventually Alex gets frustrated, stops, and toggles his availability to ``off''.
The phone rests quietly for the rest of the drive.

As Alex parks his car, Lupe's response comes in:
she'll be by in 30 minutes to start the laundry and will do wash \& dry.
Alex asks if Lupe can fold the clothing and, in the meantime while the laundry is running,
clean up his apartment. She'll later agree for an added fee.

Alex gets to Philz, opens his laptop, and navigates to his work site.
He's just received a notification: he's been granted the qualification to squash Google bugs,
so his eligible work list has expanded from around 10,000 tasks to 30,000.
This is excellent; most of the new tasks Alex has seeing lately were either too far out of his wheelhouse
or it was too unclear what the problem was; he's wary of taking on work where the input might be unclear;
once he worked on an issue that ended up taking almost a week to sort out and only paid \$10.

If Alex puts more time into squashing bugs on Facebook,
he might be promoted to a higher role that can take on more interesting,
lucrative tasks,
but the low--hanging fruit at Google and Apple are more plentiful,
so Alex opts to continue ``playing the field'' where the tasks are easier,
and more importantly more likely to pay off.

A few hours pass, and Alex has closed 40 bugs, making \$60;
he messages his friends to see where they are:

\begin{itemize}
\item Stephen is on his way, just dropping someone off before going off the clock
(\textit{Stephen drives for a living,
usually working nights and weekends in the city but occasionally picking up airport runs, like now}).
He should be by in around 15 minutes.
\item Erin messages that she has to cancel.
She's been getting jobs rejected lately
(\textit{Erin is a designer for more firms than she can count});
she needs to pick up some more gigs to make up the money she's lost from those rejections
(and, if possible, to recoup her reputation after a string of rejections).
\end{itemize}

Stephen catches up with Alex at Philz;
like Erin, Stephen has been having trouble with customers lately, but his work
requires him to go out and drive, putting himself ``in harm's way,'' as it were,
% Stephen's been having a rough go of things just as Erin has, but his work
% means that he has to get out there to do work,
whereas Erin's work is all at home.
In that sense, her work is a little more desirable;
Erin \textit{could}, she jokes, do all her work from Bermuda.
She's never been, but she's saving for it. Maybe next year.

Stephen tells Alex about some of the terrible passengers he had over the weekend:
two were absolutely trashed, both making a mess in his car;
a third passenger seemed to make it his mission to harass Stephen the whole way,
interrogating him for the route he took (even though his app navigates for him),
and generally treating him terribly.

Alex sighs and tells Stephen to just keep at it; 
Stephen sighs and says that he can put up with it as long as he doesn't get ``fat--fingered''
--- a term referencing when passengers accidentally (or deliberately) give their drivers one--star reviews.
Another of those, and Stephen might get suspended from the platform.
And it changes every day; today Palo Alto's threshold is 4.6; yesterday it was 4.3; tomorrow it might be 4.7
--- and if you wake up one morning and your rating is too low for the are you live in,
then you'll just have to do work somewhere else or stop working.

Alex glances at his phone.
The screen keeps flashing subtly as new jobs stream in.
One from Twitter.
Another, this time from Apple.
This one from Microsoft pays (almost suspiciously) well.
These are all well--paying jobs
--- a few dollars for closing a simple ticket, up to as much as a C--note for something more involved.
There are other problems associated with this work,
like maintaining confidentiality and keeping everything mentally sandboxed, 
but none of it puts him in physical danger the way on--demand driving does with Stephen;
and his work is more cut--and--dry than Erin's,
so he generally knows when he submits work whether it'll be approved or rejected.

Stephen's phone buzzes this time;
someone's looking for an SUV driver for a long trip to SoCal.
It's too good an offer to pass up.
Stephen bids farewell to Alex and quickly makes his way to the door.
Alex pulls his laptop back out and gets back to work.

% \end{counted}
% \footnotesize{word count: \thewords}

\bibliographystyle{acm} % this will cause an error if there are no citations!
\bibliography{../../../references.bib}
\end{document}