### 2016-07-15 drafts/theoryPaper/proceedings.tex

\documentclass{sigchi}
\usepackage{graphics,balance,txfonts,mathptmx,color,booktabs,color,booktabs,textcomp,microtype,ccicons}
\usepackage[pdftex]{hyperref}
\usepackage[utf8]{inputenc}
\def\plaintitle{Laying Groundwork For a Worker--Centric Peer Economy}
\def\plainauthor{Ali Alkhatib, Justin Cranshaw, Andrés Monroy-Hernandez}
\def\emptyauthor{Anonymous for Review}
\def\plainkeywords{Peer economy; Collaborative markets}
\def\plaingeneralterms{Documentation, Standardization}

\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},       pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords}, bookmarksnumbered,
  pdfstartview={FitH},          colorlinks,
  citecolor=black,              filecolor=black,          linkcolor=black,
  urlcolor=linkColor,           breaklinks=true,
}

\begin{document}

\title{\plaintitle}

\numberofauthors{4}
\author{
  \alignauthor{Ali Alkhatib                 \affaddr{Stanford University}   \affaddr{Stanford, USA}
        \email{ali.alkhatib@cs.stanford.edu}}
  \alignauthor{Justin Cranshaw              \affaddr{Microsoft Research}    \affaddr{Redmond, WA, USA}
        \email{justincr@microsoft.com}}
  \alignauthor{Andrés Monroy-Hernandez      \affaddr{Microsoft Research}    \affaddr{Redmond, WA, USA}
        \email{amh@microsoft.com}}
  \alignauthor{Michael Bernstein      \affaddr{Stanford University}    \affaddr{Stanford, USA}
        \email{msb@cs.stanford.edu}}
}
\numberofauthors{1} \author{ \alignauthor{\emptyauthor} }
\maketitle

\begin{abstract}
The ``gig economy''
--- where work is measured through discrete tasks rather than through interconnected (and often long--running) careers ---
has become a site of significant frustration among workers who have come to rely on these systems and this form of work.
This paper explores some of the sources of that stress,
and attempts to relate this seemingly nascent work style to more familiar forms of labor
--- namely, ``piece work'',
a form of work which emerged at the dawn of the 20th century.
While we find and identify parallels between ``gig work'' and the more theoretically familiar ``piece work'',
we emphasize divergences which make the conflation of these ideas problematic.

We further explore opportunities and tactics to empower workers,
informed by ethnographic fieldwork including participant--observation,
semi--structured interviews,
and participatory design.
Our observation and interviews lead us to conclude that the most intuitive mechanisms for empowering workers
--- namely, labor unions advocating collectively and counter--balancing the manager's outsize leverage ---
may both structurally and culturally fail to capture the needs of ``gig workers''.
We argue that design interventions to empower ``gig workers'' must build on the theoretically familiar,
much as ``gig work'' itself draws inspiration from ``piece work'',
but that it is crucial that interventions are designed to support,
rather than reject,
the unique cultural and system qualities of gig workers.
\end{abstract}

\category{H.5.m.}{Information Interfaces and Presentation
  (e.g. HCI)}{Miscellaneous} \category{See
  \url{http://acm.org/about/class/1998/} for the full list of ACM
  classifiers. This section is required.}{}{}

\keywords{\plainkeywords}

\section*{Dealing with Monopsonies}
\section*{Collective Bargaining}

% \input{intro.tex}
% \input{background.tex}
% \input{fieldwork_advice.tex}
% \input{design_guidance.tex}
% \input{discussion.tex}

% % REFERENCES FORMAT
% % References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{../references}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: