### 2016-07-16 drafts/groundwork/design_guidance.tex

%!TEX root = proceedings.tex
\section{Emergent Themes}
We identified several aspects of the on-demand markets we studied that seem to apply in generalizable ways.
These characteristics and issues extend beyond the more narrowly scoped issues of scheduling workers and handling payments which might seem more at the focal point of designing a labor market, although those issues too must be addressed;
we focus specifically on the social negotiations that a technological system must broker.
These guidelines are illustrated using mock-ups and a mobile application front-end developed in tandem with organization C, organization B, \& organization E, and consist of the following considerations:

\begin{enumerate} \itemsep0pt \parskip0pt
  \item Constructive feedback
  \item Assigning work fairly
  \item Managing customer expectations
  \item Protecting vulnerable workers
  \item Reconciling worker identities
  \item Assessing worker qualifications
  \item Communicating worker quality
% \item Feedback
% \item Assigning Work
% \item Customer Expectations
% \item Vulnerable Workers
% \item Worker Identity
% \item Worker Qualifications
% \item Worker Quality \& Ratings
\end{enumerate}

\subsection{1.        Constructive Feedback}
\input{designGuidance/feedback.tex}
\subsection{2.        Assigning Work Fairly}
\input{designGuidance/assigningWork.tex}
\subsection{3.        Managing Customer Expectations}
\input{designGuidance/customerExpectations.tex}
\subsection{4.        Protecting Vulnerable Workers}
\input{designGuidance/vulnerableWorkers.tex}
\subsection{5.        Reconciling Worker Identities}
\input{designGuidance/workerIdentity.tex}
\subsection{6.        Assessing Worker Qualifications}
\input{designGuidance/workerQualifications.tex}
\subsection{7.        Communicating Worker Quality}
\input{designGuidance/workerQuality.tex}