### 2016-09-08 drafts/theoryPaper/proceedings.tex

\documentclass{sigchi}
\usepackage{graphics,balance,txfonts,mathptmx,color,booktabs,color,booktabs,textcomp,microtype,ccicons}
\usepackage[pdftex]{hyperref}
\usepackage[utf8]{inputenc}
\def\plaintitle{Here's a title!}
\def\plainauthor{People here!}
\def\emptyauthor{Anonymous for Review}
\def\plainkeywords{???????????}
\def\plaingeneralterms{????}

\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},       pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords}, bookmarksnumbered,
  pdfstartview={FitH},          colorlinks,
  citecolor=black,              filecolor=black,          linkcolor=black,
  urlcolor=linkColor,           breaklinks=true,
}

\begin{document}

\title{\plaintitle}

\numberofauthors{1}
\author{
    \alignauthor{Kenji Hata, Ranjay Krishna, Li Fei--Fei, Michael S. Bernstein}\\
    \affaddr{Stanford University}\\
    \email{\{kenjihata, ranjaykrishna, feifeili, msb\}@cs.stanford.edu}
  }
\maketitle
\end{document}

% \begin{abstract}
% The ``gig economy''
% --- where work is measured through discrete tasks rather than through interconnected (and often long--running) careers ---
% has become a site of significant frustration among workers who have come to rely on these systems and this form of work.
% This paper explores some of the sources of that stress,
% and attempts to relate this seemingly nascent work style to more familiar forms of labor
% --- namely, ``piece work'',
% a form of work which emerged at the dawn of the 20th century.
% While we find and identify parallels between ``gig work'' and the more theoretically familiar ``piece work'',
% we emphasize divergences which make the conflation of these ideas problematic.

% We further explore opportunities and tactics to empower workers,
% informed by ethnographic fieldwork including participant--observation,
% semi--structured interviews,
% and participatory design.
% Our observation and interviews lead us to conclude that the most intuitive mechanisms for empowering workers
% --- namely, labor unions advocating collectively and counter--balancing the manager's outsize leverage ---
% may both structurally and culturally fail to capture the needs of ``gig workers''.
% We argue that design interventions to empower ``gig workers'' must build on the theoretically familiar,
% much as ``gig work'' itself draws inspiration from ``piece work'',
% but that it is crucial that interventions are designed to support,
% rather than reject,
% the unique cultural and system qualities of gig workers.
% \end{abstract}

% \category{H.5.m.}{Information Interfaces and Presentation
%   (e.g. HCI)}{Miscellaneous} \category{See
%   \url{http://acm.org/about/class/1998/} for the full list of ACM
%   classifiers. This section is required.}{}{}

% \keywords{\plainkeywords}

% \section*{Dealing with Monopsonies}
% \section*{Collective Bargaining}

% % \input{intro.tex}
% % \input{background.tex}
% % \input{fieldwork_advice.tex}
% % \input{design_guidance.tex}
% % \input{discussion.tex}

% % % REFERENCES FORMAT
% % % References must be the same font size as other body text.
% \bibliographystyle{SIGCHI-Reference-Format}
% \bibliography{references}

% \end{document}

% %%% Local Variables:
% %%% mode: latex
% %%% TeX-master: t
% %%% End: