### 2016-11-21 admin/IRB/performances/interview_questions/main.tex
rebuttal init

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{nameref,color}
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  % pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  % pdfauthor={\emptyauthor},
  % pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
% \setlength{\parskip}{.4em}
\pagenumbering{gobble}
\begin{document}

Interviews with performance artists under this protocol are designed to be
(at least somewhat)
open--ended, but given that we're still going in with a handful of
\textbf{a priori} notions of what we want to learn about,
we have a list of prepared questions that we plan to use to prompt
further discussion about whatever proves to be of interest to participants.
That is, while we have questions about performance artists'
\hyperref[sec:Background]{backgrounds},
\hyperref[sec:LearningCraft]{learning their craft},
\hyperref[sec:EngagingAudience]{engaging with an audience that has more direct avenues of communication},
\&
\hyperref[sec:DealingWithLaw]{negotiating the political and legal issues within online platforms}.

\begin{enumerate}

  \section*{Background}
            \label{sec:Background}
    \item How long have you been (streaming | YouTubing | podcasting | etc\dots)?
    \item What did you do before that?
    \item What got you into this?

  \section*{Learning the craft}
            \label{sec:LearningCraft}
    \item I imagine there are some skills that you have to learn to be a good
          (streamer, podcaster, etc\dots); can you talk about any of the skills
          you found you had to learn to be good at this?
    \item What sorts of things do you have to become good at
          that your audience might not realize (if anything)?

  \section*{Engaging the Audience}
            \label{sec:EngagingAudience}
    \item Do you see feedback or engagement from your audience?
          How?
    \item What avenues do people take to engage with you?
          Do you actively open and manage those channels, or
          do people simply find them?
    \item How do you respond to that kind of engagement?
          That is,
            do you prompt it?
            Encourage it?
            Keep it at arm's length?
    \item Do you encounter much criticism?
          (If so\dots)
    \item How do you deal with criticism?
    \item Is this feedback and engagement overall a good thing, or
          is it generally more trouble than the value it provides?

  \section*{Dealing with ``law'' and politics}
            \label{sec:DealingWithLaw}
    \item Do any policies or rules come up in your videos?
          To give an example, do you find yourself fending off DMCA complaints,
          or having to argue that your content isn't in violation of some rule?
    \item Do rules and policies ever change?
          (Can you describe how?)
    \item What influence do you (or others in your position) have
          (if any)
          in affecting policies?
    \item How do you think of yourself
          in relation to the platform you publish on?
    \item To be more specific,
          how much agency do you feel you have in
          changing policies that don't make sense
          (or are unreasonably applied)?
\end{enumerate}

% \printbibliography{}
\end{document}