### 2016-11-21 drafts/compensating_task_search/introduction.tex

\documentclass[main]{subfiles}
\onlyinsubfile{
  \usepackage{xr-hyper}
  \usepackage{hyperref}
  % \externaldocument{nameHere}
}

\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\section{Introduction}\label{sec:introduction}

Motivated by~\cite{crowdcollab} and the general insight that a fair amount of work
isn't compensated because it doesn't fall within the confines of narrowly defined tasks,
we address the known problem of task search, described by~\cite{taskSearch}
and studied by a number of subsequent researchers.

Workers on Mechanical Turk already spend time helping one another
(e.g. see \href{https://www.reddit.com/r/HITsWorthTurkingFor/}{/r/HITsWorthTurkingFor},
a sub--community on Reddit to assist ``Turkers'' in finding worthwhile tasks),
but their reward for such work is relatively unclear;
some prior research into hiring referrals suggests that
people who make worthwhile connections are afforded a certain amount of
social capital~\cite[see][]{Gemkow2011}.

Still, social aspects of job referral as they've been studied in the past have focused

\onlyinsubfile{
  % \balance{}
  \printbibliography
}

\end{document}