### 2016-11-30 admin/IRB/performances/recruitment/main.tex

\documentclass[11pt]{article}
\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage[inline]{enumitem}
\usepackage[tiny,compact]{titlesec}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\usepackage{nameref,color}
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  % pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  % pdfauthor={\emptyauthor},
  % pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=blue,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
% \setlength{\parskip}{.4em}
\pagenumbering{gobble}
\begin{document}
Hi,

My name's Ali Alkhatib. % link to homepage
I'm a PhD student at Stanford.
I'm studying professional online performers
(like YouTubers, Twitch streamers, etc\dots)
--- basically anyone who does it as a major source of income.

Specifically, I'm looking into
the relationships people have with their platforms
--- so for instance, what it's like for
YouTubers to make their living from the videos they put on YouTube,
talking to and working with YouTube's admins,
and dealing with controversial ``meta'' topics like
sudden changes to rules on the platform.

I'm emailing because I saw your
[\textbf{online streaming/broadcasting channel}]
and I thought your perspective would be interesting.
I was wondering if you'd be willing to chat with me about anything relating to
[\textbf{streaming|YouTubing|etc\dots}]. % make it clear this'll be anonymous if desired

% What I had in mind is a chat that would probably take about 30 minutes,
% possibly with a few follow up conversations.
% I was figuring to video chat via
% FaceTime or Google Hangout or something for the first
% ``long'' conversation, and email for later follow ups.

If you're open to chatting about online performances and stuff, feel free to
\href{mailto:ali.alkhatib@cs.stanford.edu}{email me}
(\texttt{ali\allowbreak .alkhatib\allowbreak @cs\allowbreak .stanford.edu}) or
call me at my cell phone: ${+1~(408)~348-2720}$

PS: if you're curious what sort of research I generally do,
feel free to poke around at \href{https://ali-alkhatib.com}{my personal site}
(\texttt{https://\allowbreak ali-\allowbreak alkhatib\allowbreak .com}).

\end{document}