### 2016-12-05 drafts/aggregator/introduction.tex
proposal (why didn't i put this up sooner?)

\documentclass[main]{subfiles}

\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}

\section{Introduction}\label{sec:introduction}

Online labor markets ranging a variety of industries of (largely) service work
--- e.g. on--demand drivers (Uber, Lyft, Sidecar, etc\dots) and
    information work (Amazon Mechanical Turk, Upwork, CrowdFlower, etc\dots) ---
have struggled for some time with pernicious, in some cases \textit{growing},
difficulty assessing the quality of its workforce,
especially as it applies to a given task.
In other words, whether a worker on Amazon Mechanical Turk (AMT) is likely
to complete a task ``accurately'' remains an open question,
leading researchers to address this pervasive problem in a number of ways
\cite

% Labor market platforms ranging industries and styles of work
% (e.g. Uber, TaskRabbit, UpWork, and Amazon Mechanical Turk)
% have struggled for years with persistent
% (generally growing)
% challenges relating to worker qualifications.
% These issues range a wide spectrum:
% in some cases, worker qualifications are non--transferable,
% leading requesters to ``re--invent the wheel'' as they attempt to determine
% in their own way whether a potential worker is qualified and reliable.

% Problems determining a worker's qualifications start on day 1;
% labor markets begin their relationships with new workers
% almost entirely uncertain about the worker's competence in any type of task.
% Gathering this information through qualification exams is generally
% time--consuming and costly.

% Challenges mount as workers' skill sets develop;
% work requiring more training and skill (for example, translating or programming)
% are either verified by individual \textit{requesters} (e.g. Amazon Mechanical Turk)
% or are verified by the \textit{platform} itself (e.g. UpWork).
% While the \textit{UpWork} model avoids needless repetitive work by
% generally consolidating qualifying exams at the platform level,
% these labor platforms nevertheless find themselves in the unenviable
% (and often unexpected) position of
% having to develop new qualifications exams to outpace would--be cheaters.

\onlyinsubfile{
% \balance{}
\printbibliography{}
}

\end{document}