### 2016-12-08 drafts/labor/trackingWork.tex

\documentclass[chi_draft]{sigchi}

\usepackage{xr-hyper}
\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes,cleveref,textcase}
\usepackage[normalem]{ulem}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\usepackage[T1]{fontenc}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{mfirstuc,lastpage}
\usepackage[en-GB]{datetime2}
\usepackage{fancyhdr}
\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage[utf8]{inputenc} % for a UTF8 editor only

% \bibliographystyle{SIGCHI-Reference-Format}
\renewcommand*{\notesname}{Graveyard of old paragraphs}
\makeatletter
\def\blx@maxline{77}
\makeatother

\providecommand*{\namerefl}[1]{%
\refused{#1}%
\MakeTextLowercase{%
\getrefbykeydefault{#1}{name}{??}%
}%
}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.

\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
   % label=\alph*),
 label=\arabic*),
}

\setlist{nolistsep}
\def\plaintitle{Examining Crowd Work Through The Historical Lens of Piecework}
\def\plainauthor{Ali Alkhatib, Margaret Levi, Michael Scott Bernstein}
\def\emptyauthor{}
\def\plainkeywords{Crowdsourcing; on--demand labor; gig work}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}
\definecolor{PineGreen}{HTML}{008800}
\definecolor{BrickRed}{HTML}{FF0000}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{BrickRed}[al2: #1]}}
\newcommand{\TO}{\textit{Turkopticon}}
\newcommand{\DO}{\textit{Dynamo}}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
  pdfauthor={\plainauthor},
  % pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\bibliography{references}

\newcommand{\subsubsubsection}[1]{\textit{#1}}
\newcommand{\pieceworkpers}{Piecework's perspective}
\newcommand{\crowdworkpers}{Crowd work's perspective}
\newcommand{\whatchanged}{Comparing the phenomena}
\newcommand{\implication}{Implications for crowd work}
\newcommand{\pieceworkdefinition}{piecework is the paradigm of renumeration that is largely made compelling for its use of
``payment for results'',
leading to and leveraging narrower skill sets required for narrowly defined tasks,
and affording workers some amount of freedom to complete tasks
(and consequently earn money)
at at whatever rate and in whichever manner they wish} % do i use this more than once? even that much?

\newcommand{\onlyinsubfile}[1]{#1}
\newcommand{\notinsubfile}[1]{}
\newcommand{\note}[1]{\footnote{#1}}
\newcommand{\topic}[1]{\textbf{#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
\newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!

\let\internalNoteali\ali
\let\internalNotemsb\msb
\onlyinsubfile{\renewcommand{\ali}[1]{\onlyinsubfile{\internalNoteali{#1}}}}
\onlyinsubfile{\renewcommand{\msb}[1]{\onlyinsubfile{\internalNotemsb{#1}}}}

\renewcommand{\headrulewidth}{0pt}

% \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
\renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)
\begin{document}
\renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
\renewcommand{\joke}[1]{} % don't do anything with \jokes! Just throw them away! (Better to drop it than accidentally include it)
\renewcommand{\onlyinsubfile}[1]{}
\renewcommand{\notinsubfile}[1]{#1}
\renewcommand{\note}[1]{\onlyinsubfile{\note{#1}}}

\title{\plaintitle}

\numberofauthors{3}
\author{%
  \alignauthor{Ali Alkhatib\\
    \affaddr{Stanford University}\\
    \affaddr{Stanford, USA}\\
    \email{ali.alkhatib@cs.stanford.edu}}\\
  \alignauthor{Michael Scott Bernstein\\
    \affaddr{Stanford University}\\
    \affaddr{Stanford, USA}\\
    \email{msb@cs.stanford.edu}}\\
  \alignauthor{Margaret Levi\\
    \affaddr{CASBS}\\
    \affaddr{Stanford, USA}\\
    \email{mlevi@stanford.edu}}\\
}
% \author{
% \alignauthor{Anonymized for Review\\
% \affaddr{~}\\
%     \affaddr{~}\\
%     \email{~}}\\
% }
\maketitle
\subfile{abstract.tex}

\category{H.5.3.}{Information Interfaces and Presentation
  (e.g. HCI)}{Group and Organization Interfaces}

\keywords{\plainkeywords}
\subfile{introduction.tex}
\subfile{piecework_lit.tex}
\subfile{cases.tex}
\subfile{discussion.tex}
\subfile{conclusion.tex}

\balance{}
\printbibliography

\clearpage
\nobalance{}
\endnotetext{
}

\end{document}