### 2017-01-08 drafts/labor/cases.tex
rough pw perspective

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother
\begin{document}
\section{Research Questions}

{Research in crowdsourcing has spent the better part of a decade
exploring how to grow the limits of crowdsourcing,
finding the boundaries of crowd work and microtasks.}
This has largely involved iteratively
identifying barriers on complexity and
overcoming them through novel designs of work--flows and processes (e.g.~\cite{bernsteinSoylent,foundry,crowdForgeKittur}).
The question then
has become \textit{whether} there are limits to crowdsourcing,
and if so, what factors determine them.
To this question, a number of contributions to the field have pressed for answers.

The exploration of on--demand work's potential and limits has principally looked at
manipulating and extending along three dimensions:
First, \namerefl{sec:complexity}.
Second, \namerefl{sec:decomposition}.
And third, \namerefl{sec:relationships}.
We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
decomposed,  contextually abstracted, and
made more resilient to attrition of various forms.
We'll also point to corresponding piecework literature addressing these aspects.
% \msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
Finally, we'll discuss how these elements will serve
to constrain the upper and lower bounds of crowdsourcing as it relates
to the question of the furthest limits of crowdsourcing.

\notinsubfile{
  \subfile{complexity.tex}
  \subfile{decomposition.tex}
  \subfile{relationships.tex}
}
\onlyinsubfile{
  \printbibliography{}
}

\end{document}