### 2017-01-08 drafts/labor/conclusion.tex

\documentclass[trackingWork]{subfiles}
\makeatletter
\def\blx@maxline{77}
\makeatother

\onlyinsubfile{
\usepackage{xr-hyper}
\usepackage{hyperref}
\externaldocument{complexity}
\externaldocument{decomposition}
\externaldocument{relationships}
}

\begin{document}
\section{Conclusion}
On--demand work is not new, but a contemporary instantiations of piecework.
In this paper, we reconsider three major research questions in on--demand work using the lens of piecework:
\begin{inlinelist}
  \item ``\namerefl{sec:complexity}?'';
  \item ``\namerefl{sec:decomposition}?''; and
  \item ``\namerefl{sec:relationships}?''
\end{inlinelist}
To do so, we draw on piecework scholarship to inform analyses of what has changed, what hasn't, and may change change soon.
Reciprocally, we believe that modern on--demand work will teach us about the broader phenomenon of piecework as well.
If history really does repeat itself, the best we can do is be prepared.

\end{document}