### 2017-01-11 drafts/labor/trackingWork.tex

\documentclass{sigchi}

\usepackage{xr-hyper}
\usepackage{todonotes,txfonts,balance,graphics,color,alphalph,subfiles,appendix,endnotes}
\usepackage{booktabs,textcomp,microtype,ccicons,enumitem,moreenum,nth,endnotes,cleveref,textcase}
\usepackage[normalem]{ulem}
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\usepackage[T1]{fontenc}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{mfirstuc,lastpage}
\usepackage[en-GB]{datetime2}
\usepackage{fancyhdr,tabularx}
\usepackage[all]{hypcap}  % Fixes bug in hyperref caption linking
\usepackage[utf8]{inputenc} % for a UTF8 editor only

% \bibliographystyle{SIGCHI-Reference-Format}
\renewcommand*{\notesname}{Graveyard of old paragraphs}
\makeatletter
\def\blx@maxline{77}
\makeatother

\providecommand*{\namerefl}[1]{%
\refused{#1}%
\MakeTextLowercase{%
\getrefbykeydefault{#1}{name}{??}%
}%
}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.

\newlist{numberlist}{enumerate*}{1}
\setlist*[numberlist,1]{%
   % label=\alph*),
 label={\nthwords*,},
}

\newlist{Numberlist}{enumerate*}{1}
\setlist*[Numberlist,1]{%
   % label=\alph*),
 label={\Nthwords*,},
}
\newlist{inlinelist}{enumerate*}{1}
\setlist*[inlinelist,1]{%
   % label=\alph*),
 label=\arabic*),
}

\setlist{nolistsep}
\def\plaintitle{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
% \def\plaintitle{Crowd Work and Gig Work [as] Seen Through the Historical Lens of Piecework}
\def\plainauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi}
\def\emptyauthor{}
\def\plainkeywords{Crowd work; gig work; on--demand work; piecework}
\def\plaingeneralterms{Group and Organization Interfaces}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}
\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}
\newcommand{\TO}{\textit{Turkopticon}}
\newcommand{\DO}{\textit{Dynamo}}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
  pdfauthor={\plainauthor},
  % pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\bibliography{references}

\newcommand{\subsubsubsection}[1]{\textit{#1}}
\newcommand{\pieceworkpers}{The perspective of piecework}
\newcommand{\crowdworkpers}{The perspective of on--demand work}
\newcommand{\whatchanged}{Comparing the phenomena}
\newcommand{\implication}{Implications for on--demand work}
\newcommand{\pieceworkdefinition}{piecework is the paradigm of remuneration that is largely made compelling for its use of
``payment for results'',
leading to and leveraging narrower skill sets required for narrowly defined tasks,
and affording workers some amount of freedom to complete tasks
(and consequently earn money)
at at whatever rate and in whichever manner they wish} % do i use this more than once? even that much?

\newcommand{\onlyinsubfile}[1]{#1}
\newcommand{\notinsubfile}[1]{}
\newcommand{\note}[1]{\footnote{#1}}
\newcommand{\topic}[1]{{\color{Blue}#1}} % if you're in a subfile, bold the \topic sentences to make it easier to skim/diagnose logical issues
\newcommand{\joke}[1]{{\color{JokeGreen}#1}} % if you're in a subfile, make the joke JokeGreen colored so people know I'm being funny!

\let\internalNoteali\ali
\let\internalNotemsb\msb
\onlyinsubfile{\renewcommand{\ali}[1]{\onlyinsubfile{\internalNoteali{#1}}}}
\onlyinsubfile{\renewcommand{\msb}[1]{\onlyinsubfile{\internalNotemsb{#1}}}}

\renewcommand{\headrulewidth}{0pt}

% \renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
\renewcommand{\footnotesize}{\normalsize} % i don't need small text! (no but really remove this before submitting)

\begin{document}
\CopyrightYear{2017} 
\setcopyright{acmlicensed}
\conferenceinfo{CHI 2017,}{May 06 - 11, 2017, Denver, CO, USA}
\isbn{978-1-4503-4655-9/17/05}\acmPrice{\$15.00}
\doi{http://dx.doi.org/10.1145/3025453.3025974}
\renewcommand{\topic}[1]{#1} % don't do anything with \topic sentences; this is just for drafting 
\renewcommand{\joke}[1]{} % don't do anything with \jokes! Just throw them away! (Better to drop it than accidentally include it)
\renewcommand{\onlyinsubfile}[1]{}
\renewcommand{\notinsubfile}[1]{#1}
\renewcommand{\note}[1]{\onlyinsubfile{\note{#1}}}

\title{\plaintitle}

\numberofauthors{1}
\author{%
  \alignauthor{Ali Alkhatib, Michael S. Bernstein, Margaret Levi\\
    \affaddr{Computer Science Department and CASBS}\\
    \affaddr{Stanford University}\\
    \email{\{ali.alkhatib, msb\}@cs.stanford.edu, mlevi@stanford.edu}}\\
}
% \author{
% \alignauthor{Anonymized for Review\\
% \affaddr{~}\\
%     \affaddr{~}\\
%     \email{~}}\\
% }
\maketitle
\subfile{abstract.tex}

\category{H.5.3.}{Information Interfaces and Presentation
  (e.g. HCI)}{Group and Organization Interfaces}

\keywords{\plainkeywords}
\ali{maybe put a table with high level summaries}
\subfile{introduction.tex}
\subfile{piecework_lit.tex}
\subfile{cases.tex}
\subfile{discussion.tex}
\subfile{conclusion.tex}

% \balance{}
% \printbibliography

\end{document}