### 2017-05-03 chi/piecework/presentation/introduction.tex
complexity lit review done?

\documentclass[presentation]{subfiles}

\onlyinsubfile{
  % \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
  % \bibliography{~/Library/texmf/bibtex/bib/local/references}
  % \beamertemplategridbackground[1mm]
}

\begin{document}
% \section*{Introduction}

\begin{frame}{Open problems in crowdsourcing}
\begin{columns}
\begin{column}[T]{0.5\textwidth}
  \begin{itemize}%[<+>]
    \item<2,4> \textbf{\large Complexity}

    \scriptsize{
    \textcite{suzukiAtelier,KimStoria,yuanAlmost,Yu2016b,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
    }
    \vspace{8mm}

    \item<3> \textbf{\large Decomposition}

    \scriptsize{
    \textcite{sensitiveTasks,LykourentzouPersonalityMatters,Law:2016:CKC:2858036.2858144,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
    }
    \vspace{8mm}

    \item<5> \textbf{\large Relationships}

    \scriptsize{
      \textcite{turkopticon,storiesIraniSilberman,crowdcollab,takingAHITMcInnis,dynamo,uberAlgorithm}
    }

  \end{itemize}

\end{column}

\begin{column}[T]{0.25\textwidth}
  \begin{figure}
    \centering
    \vspace{-5mm}

    \visible<2->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}
    }\visible<4->{\includegraphics[width=.33\textwidth]{figures/complexity/paper_resized.png}
                  \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
    }

    \vspace{5mm}
    \includegraphics[width=\textwidth]{figures/complexity/hexRepeat.png}
    \vspace{5mm}

    \visible<3->{
    \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
    }

    \visible<5->{
      \vspace{10mm}
      \includegraphics[width=\textwidth]{figures/idk.png}
    }
  \end{figure}

\end{column}

\begin{column}[T]{0.25\textwidth}
\centering
\vspace{5mm}

\visible<2->{
  \small{Complexity}
  \vspace{10mm}
}

\small{Tasks}

\visible<3->{
  \vspace{10mm}
  \small{Decomposition}
}

\end{column}
\end{columns}
\end{frame}

\begin{frame}[standout]
    What is the future of work?
\end{frame}

\begin{frame}{What is the future of work?}
    How will \textbf{technology} affect the \textbf{complexity} of the work that on--demand workers do?

    What are the \textbf{limits} of complexity in on--demand work?

    How can we \textbf{reach} those limits?
\end{frame}

\notinsubfile{
  \subfile{timeline.tex} % this takes a while to compile...
}

\begin{frame}{Introduction}
  We hope to provide:
      \begin{itemize}
        \item A useful ontological lens for making sense of crowdsourcing and gig work (which we collectively call ``\textit{on--demand work}'') as a resurgence of \textit{piecework}.
        \item A method for making sense of contemporary phenomena through \textit{historical analysis}.
      \end{itemize}
\end{frame}

\end{document}