### 2017-05-09 chi/piecework/presentation/presentation.tex

\documentclass{beamer}

\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer,nth,xcolor,svg,subfiles,chronosys,cancel}
\usepackage{pgfcalendar}
\usepackage{datenumber,xparse,fp}

\usetikzlibrary{arrows,automata}
\usepackage[citestyle=authoryear-comp,backend=bibtex]{biblatex}
\usepackage[export]{adjustbox}
\usepackage[en-US]{datetime2}

\bibliography{references}
\usetheme{metropolis}
\beamertemplatenavigationsymbolsempty
\definecolor{stanfordRed}{HTML}{8C1515}
\definecolor{metroBackground}{RGB}{250,250,250}

% Recent Notes from msb
% [x] I suggest that you come out with a one sentence summary of your main thesis after 2 and before 3. “What I’m going to argue in this talk is that on-demand work is a modern instantiation of a much older phenomenon, piecework [put the word on the slide], and that we’re going to be able to use  the history of piecework over the past two hundred years to shed light on several unanswered questions about the unfolding phenomenon of on-demand work.”
% [x] This way people know what you’re winding up for in this talk
% [x] don’t get rid of slide 5. I’m just saying put one framing sentence up front here.
% [x] 3: these are not the canonical cites for complexity. kittur crowdforge is most famous
% [-] 3: put the task hexagons at the top, not in between. the hierarchy I see what you’re trying to do, but it’s not really working
% [x] When you introduce the piecework term, give several examples (with pictures) that listeners will be familiar with.
% [x] You never stopped to equate piecework to on demand work. Why does piecework = on demand work? (in particular, because on demand work ALSO pays people per task)
% [x] 3: your complexity example is being built out of triangles, not hexagons
% [x] Get rid of the citation numbers, and replace it with years Alkhatib 2017, not Alkhatib [1]
% [>] 3: I think this isn’t focused correctly. I’d get rid of this slide and ONLY motivate the complexity question. Then mention briefly that in the paper you also talk about decomposition and worker relationships, but you need 10 pages and 166 references rather than a 15 minute talk to do all that. Return to these other themes at the end of the talk and mention some short take aways from each one.
% [x] 4: doesn’t make sense to ask “how will we reach those limits?”
% [?] 7: the top timeline is events, the bottom is citations. change the bottom one to be events
% [?] 8: this is WAY too late to still be doing an intro. You’re five minutes into the talk. Intro should be done by 2min.
% [x] 12: the details here are irrelevant to complexity. why does it matter that they could be fired at will, overworked, underpaid? shouldn’t we be focusing on how piecework enabled the complexity? If it’s not important (as you say), then don’t include it. Parsimony, yay!
% [x] 12: another example of a citation I can’t follow by watching slides
% [x] 14: ditch the shrugging guy, he’s really weird
% [x] 15: trains — this has nothing to do with complexity…? cut?
%      revising so the mini-thesis is clearer; what did each of these cases contribute?
% [x] 15: fordism — again, explain what this has to do with complexity? what does fordism have to do with complex outcomes? spell it out. what should we be learning from this?
% [x] 16: technology makes complex tasks relatively trivial — this is an overly broad claim. makes SOME?
% [?] 17: speed up The Knowledge/Gmaps example. I think it can go twice as fast and still carry as much weight
% will try this out in practice talk

\setbeamertemplate{frametitle}
{\vskip-3pt
  \leavevmode
  \hbox{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=5.8ex,dp=1ex]{frametitle}%
    \raggedright\hspace*{1em}\large{\insertframetitle}
  \end{beamercolorbox}
  }%
}
\setbeamertemplate{caption}{\raggedright\footnotesize{\insertcaption}\par}

\addtobeamertemplate{frametitle}{}{%
    \begin{textblock*}{28mm}(\paperwidth-40mm,-12mm)
      \includegraphics[height=10mm]{figures/origami-1.png}
      \includegraphics[height=10mm]{figures/origami-2.png}
      \includegraphics[height=8mm]{figures/origami-white.png}
    \end{textblock*}}

\input{timeline_code}

\DeclareCiteCommand{\eventcite}{}{\chronosevent{\thefield{year}}{\printnames{labelname}}}{}{}

\definecolor{Orange}{HTML}{E98125}
\definecolor{Blue}{HTML}{0645AD}
\hypersetup{colorlinks,linkcolor=,urlcolor=Blue}

\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}

\newenvironment{mystepwiseitemize}{\begin{itemize}[<+-| alert@+>]}{\end{itemize}}
% Theme colors are derived from these two elements
\setbeamercolor{alerted text}{fg=Orange}
\setsansfont[BoldFont={Source Sans Pro Semibold},
              Numbers={OldStyle}]{Source Sans Pro}
\setmonofont{Source Code Pro}
\setbeamercolor{institute in head/foot}{fg=stanfordRed}
\setbeamercovered{transparent}
\setbeamercovered{again covered={\opaqueness<1->{15}}}

\title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
% \metroset{titleformat=smallcaps}

\author{\textbf{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
\texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}\normalsize{} }}

\institute[Stanford]{Stanford University}
\date{\today}

\hypersetup{
  colorlinks = true,
  % linkcolor = blue,
  citecolor = blue
}

\newcommand{\onlyinsubfile}[1]{#1}
\newcommand{\notinsubfile}[1]{}
\begin{document}
\renewcommand{\onlyinsubfile}[1]{}
\renewcommand{\notinsubfile}[1]{#1}

\begin{frame}
\titlepage
\end{frame}
% \setbeamercovered{transparent}

\subfile{introduction}
\subfile{complexity}
\subfile{closing}
\subfile{contact}

% \begin{frame}[allowframebreaks]{References}
% \printbibliography{}
% \end{frame}

\end{document}