### 2017-05-10 chi/piecework/presentation/presentation.tex

\documentclass[aspectratio=169,12pt]{beamer} % or 14 or 17 or 20

\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer}
\usepackage{pgfcalendar,pgfornament,svg,subfiles,chronosys,cancel,xcolor,color,nth}
\usepackage[outline]{contour}
\usepackage{datenumber,xparse,fp}
\usetikzlibrary{arrows,automata}
\usepackage[citestyle=authoryear-comp,backend=bibtex]{biblatex}
\usepackage[export]{adjustbox}
\usepackage[en-US]{datetime2}
\bibliography{references}
\usetheme{metropolis}
% \linespread{1}
% \usecolortheme{metropolis}
\beamertemplatenavigationsymbolsempty
\definecolor{stanfordRed}{HTML}{8C1515}
\definecolor{metroBackground}{RGB}{250,250,250}

% notes from last practice talk (5/9 morning):
% [x] spacing in the definitions — piecework should be more separated from the other definitions to make clearer that it's different
%      tried something different. now it's Source Serif Pro. Thoughts?
% [x] space the crowdwork and gigwork images (uber and AMT) further down to separate them from the piecework images
% [x] move the timeline to after the historical analysis slide
% [x] return to the standout slide at the end as a way of wrapping up
% [x] on that note; instead of the slide that just says "complexity", maybe take the slide with the three panels (comp, decomp, relationships) and put that in its place, and then have a transition where i gray out decomp and relationships. then at the end be like "there's a lot more in the paper, like decomposition and relationships, but we don't have time to get deeply into them. i encourage you to read the paper if you found this talk provocative [or something], as we follow a similar style for the other discussions"
% [x] there's a flickering thing that goes on in the pdf viewer i used that is super obnoxious. some other pdf viewers are less bad about it. i'm gonna trial a few viewers since i just have to go to fullscreen and click next

% [ ] slide 8: title="introduction" not good. i think i need to move all of this stuff around so this slide is earlier and i can arguably say that the timeline (slide 7) is getting into the body of the talk (all of the events correspond to advances in *complexity*, not just crowd/piecework) 
% [ ] slide 10: i can give specific examples; "soylent _revised_ academic papers; crowd research _generates new papers_" is kind of salient 
% [x] also slide 10: "from whole cloth" is both idiomatic and possibly the wrong idiom, so i'm just going to change it to something like "completely anew" or something 
% [ ] generally: need to figure out whether to keep the origami birds. they look nice but the reference to breaking tasks down into triangles and those triangles making it possible to make more complex, nuanced things is totally lost. it feels more like an inside joke that the audience won't get 
% [x] reference/explain the chart on slide 13, because i don't really explicitly reference it. it's just an image showing several people doing work, contributing to and verifying each other's work 
% [x] slide 14: farm workers need an image (mentioned before) 
% [ ] slide 16 felt super confused and foggy. maybe it was just too abrupt. i'll try to practice my transition to it 
% [ ] slide 17: i'm dropping figure B and maybe the tiny image on the left. sound okay? 
% [ ] slide 18: needs real content and not just internal notes 
% [x] slide 20: i actually left out the relationships panel. also no visualizations. thoughts? 
% [x] slide 21: it's not twofold if i have 3 things -_- 
% [ ] slide 22: "thanks and questions?"

% more of MSB's notes
% !!!!!! [x] spend a little more time motivating why the complexity question is important. If we knew the answer to this, we would be able to predict what kind of worker displacements might happen, build appropriate infrastructure, create strong worker advocate groups, etc. 
% !!!!!! [x] 11: I think this needs to be bounded. you need to argue that in fact NOT ALL complex stuff has been achieved. e.g., the three examples you gave, while we’ve made progress, are basically still open questions. and are they fundamentally impossible, or just facing a lack of engineering imagination? that’s what we turn to piecework to answer. <— important transition, it frames the question
% !!!!!! [x] I need to hear the argument that you’re going to make about what piecework has to say about complexity, before you dive into the “ hat does piecework say?" history. otherwise it feels like a wandering tour rather than an argument.
% !!!!!! [ ] Slide 20: instead of just handwaving that we talk about it, summarize the big takeaways from your Table 1. e.g., "Decomposition: After scientific management matured, piecework began specialized training to create experts in narrow tasks. A similar shift seems feasible with on–demand work.” and “Relationships: While worker frustrations are similar, the decentralized nature of on–demand work will limit collective action until there exist platforms to coordinate and exert pressure.”
% [x] Don’t introduce your name and title; the session chair will do that for you. Make sure to show up to the room in the break before your session (at least 15min early) to test your laptop and introduce yourself to the session chair
% [x] 2: Upwork, not UpWork
% [x] 2: top bullet is services, bottom is categories. make them consistent
% [x] Intro is way better now. congrats!
% [ ] On final slide, thank Stanford Cyber Initiative for funding
% [ ] First half of Slide 8 seems redundant. Cut? And just focus on the second half.
% [ ] 10: wrap this together — essentially you need to say “so this reduces to a question of what can, and what can’t be crowdsourced. What are the limits, if any?” and put something like that on the slide
% [x] 11: inconsistent language use in titles, for example “crowdsorucing” instead of “on-demand work” in the title? unless this is intentional
% [ ] (by the time you get to automobiles, I don’t know why you’re telling me what you’re telling me)
% [ ] 13: not clear how ascention and descention of stars is “in the process”. I suggest pointing out instead that he reduced it to simple addition etc calculations that people passed between each other
% [ ] I don’t know what “marginal complexity” means. Either explain it or change the term
% [ ] 15: change “values” to “payment”…clearer
% [ ] 16 or after: make clear that technology hasn’t fundamentally changed the relationship. we can still do the equivalent of building planes, and still can’t do the equivalent of fixing planes (things that require breadth of knowledge and flexibility)
% [x] 17: don’t spend time laying out the dual hypotheses. just say that clearly this is either causing changes in the brain, or selecting for a small subset of people (in one sentence) — either way not a complex task that just anyone can do
% [ ] 21: be clear on the slide bullet three that this is about History as a lens on contemporary phenomena. right now it’s not clear

\setbeamertemplate{frametitle}
{\vskip-3pt
  \leavevmode
  \hbox{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=5.8ex,dp=1ex]{frametitle}%
    \raggedright\hspace*{1em}\large{\insertframetitle}
  \end{beamercolorbox}
  }%
}
\setbeamertemplate{caption}{\raggedright\footnotesize{\insertcaption}\par}

\addtobeamertemplate{frametitle}{}{%
    \begin{textblock*}{28mm}(\paperwidth-40mm,-12mm)
      \includegraphics[height=10mm]{figures/origami-1.png}
      \includegraphics[height=10mm]{figures/origami-2.png}
      \includegraphics[height=8mm]{figures/origami-white.png}
    \end{textblock*}}

\input{timeline_code}

\DeclareCiteCommand{\eventcite}{}{\chronosevent{\thefield{year}}{\printnames{labelname}}}{}{}

\definecolor{Orange}{HTML}{E98125}
\definecolor{Blue}{HTML}{0645AD}
\hypersetup{colorlinks,linkcolor=,urlcolor=Blue}

\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}

\newenvironment{mystepwiseitemize}{\begin{itemize}[<+-| alert@+>]}{\end{itemize}}
% Theme colors are derived from these two elements
\setbeamercolor{alerted text}{fg=Orange}
\setsansfont[BoldFont={Source Sans Pro Bold},
              Numbers={OldStyle}]{Source Sans Pro}
\setmainfont[BoldFont={Source Serif Pro Semibold},
              Numbers={OldStyle}]{Source Serif Pro}
\setmonofont{Source Code Pro}
\setbeamercolor{institute in head/foot}{fg=stanfordRed}
\setbeamercovered{transparent}
\setbeamercovered{again covered={\opaqueness<1->{15}}}

\title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
% \metroset{titleformat=smallcaps}
\contourlength{0.1pt}
\contournumber{10}
\author{\textbf{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
\texttt{ \scriptsize{\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}\normalsize{} }}

\institute[Stanford]{Stanford University}
\date{\today}

\hypersetup{
  colorlinks = true,
  % linkcolor = blue,
  citecolor = blue
}

\newcommand{\onlyinsubfile}[1]{#1}
\newcommand{\notinsubfile}[1]{}
\begin{document}
\renewcommand{\onlyinsubfile}[1]{}
\renewcommand{\notinsubfile}[1]{#1}

\begin{frame}
\titlepage
\end{frame}
% \setbeamercovered{transparent}

\subfile{introduction}
\subfile{complexity}
\subfile{closing}
\againframe[standout]{takeaway}
\subfile{contact}

\begin{frame}[allowframebreaks]{References}
  \printbibliography{}
\end{frame}

\end{document}