### 2017-05-12 chi/piecework/presentation/presentation.tex
fixed talk date to actual given date!

\documentclass[aspectratio=169,12pt,noslidenumbers]{beamer} % or 14 or 17 or 20

\usepackage{tikz,pgf,lmodern,textpos,hyperref,graphicx,booktabs,appendixnumberbeamer}
\usepackage{pgfcalendar,pgfornament,svg,subfiles,chronosys,cancel,xcolor,color,nth}
\usepackage[outline]{contour}
\usepackage{datenumber,xparse,fp}
\usetikzlibrary{arrows,automata}
\usepackage[citestyle=authoryear-comp,backend=bibtex]{biblatex}
\usepackage[export]{adjustbox}
\usepackage[en-US]{datetime2}
\bibliography{references}
\usetheme{metropolis}
\beamertemplatenavigationsymbolsempty
\definecolor{stanfordRed}{HTML}{8C1515}
\definecolor{metroBackground}{RGB}{250,250,250}

\setbeamertemplate{frametitle}
{\vskip-3pt
  \leavevmode
  \hbox{%
  \begin{beamercolorbox}[wd=\paperwidth,ht=5.8ex,dp=1ex]{frametitle}%
    \raggedright\hspace*{1em}\large{\insertframetitle}
  \end{beamercolorbox}
  }%
}
\setbeamertemplate{caption}{\raggedright\footnotesize{\insertcaption}\par}

\addtobeamertemplate{frametitle}{}{%
    \begin{textblock*}{28mm}(\paperwidth-40mm,-12mm)
      \includegraphics[height=10mm]{figures/origami-1.png}
      \includegraphics[height=10mm]{figures/origami-2.png}
      \includegraphics[height=8mm]{figures/origami-white.png}
    \end{textblock*}}

\input{timeline_code}

\DeclareCiteCommand{\eventcite}{}{\chronosevent{\thefield{year}}{\printnames{labelname}}}{}{}

\definecolor{Orange}{HTML}{E98125}
\definecolor{Blue}{HTML}{0645AD}
\hypersetup{colorlinks,linkcolor=,urlcolor=Blue}

\definecolor{PineGreen}{HTML}{008800}
\definecolor{Red}{HTML}{FF0000}
\definecolor{Blue}{HTML}{0000FF}
\definecolor{JokeGreen}{HTML}{00C953}
\newcommand{\msb}[1]{{\color{PineGreen}[MSB: #1]}}
\newcommand{\ali}[1]{{\color{Red}[al2: #1]}}

\newenvironment{mystepwiseitemize}{\begin{itemize}[<+-| alert@+>]}{\end{itemize}}
% Theme colors are derived from these two elements
\setbeamercolor{alerted text}{fg=Orange}
\setsansfont[BoldFont={Source Sans Pro Bold},
              Numbers={OldStyle}]{Source Sans Pro}
\setmainfont[BoldFont={Source Serif Pro Semibold},
              Numbers={OldStyle}]{Source Serif Pro}
\setmonofont{Source Code Pro}
\setbeamercolor{institute in head/foot}{fg=stanfordRed}
\setbeamercovered{transparent}
\setbeamercovered{again covered={\opaqueness<1->{15}}}

\title{Examining Crowd Work and Gig Work Through The Historical Lens of Piecework}
% \metroset{titleformat=smallcaps}
\contourlength{0.1pt}
\contournumber{10}
\author{{Ali Alkhatib}, Michael Bernstein, Margaret Levi\\
\href{mailto:ali.alkhatib@cs.stanford.edu}{ali.alkhatib@cs.stanford.edu} ||
         \href{http://twitter.com/_alialkhatib}{@\_alialkhatib}}

\institute[Stanford]{Stanford University}
% \newdate{talkDate}{10}{05}{2017}
\date{May 10, 2017}
% \date{\talkDate}

\hypersetup{
  colorlinks = true,
  % linkcolor = blue,
  citecolor = blue
}

\newcommand{\onlyinsubfile}[1]{#1}
\newcommand{\notinsubfile}[1]{}
\begin{document}
\renewcommand{\onlyinsubfile}[1]{}
\renewcommand{\notinsubfile}[1]{#1}

\begin{frame}
\titlepage
\end{frame}
% \setbeamercovered{transparent}

\subfile{introduction}
\subfile{complexity}
\subfile{closing}
\againframe[standout]{takeaway}
\subfile{contact}

\begin{frame}[allowframebreaks]{References}
  \printbibliography{}
\end{frame}

\end{document}