### 2017-07-09 admin/IRB/vitae/consent/main.tex
added acmart (and other stuff?)

\documentclass[10pt]{article}
\usepackage{balance,graphics,setspace,parskip,sourceserifpro,textcomp,titling,nameref,color,fancyhdr,paralist}
\usepackage[T1]{fontenc}
\usepackage[margin=1in,headheight=14pt]{geometry}
\usepackage[inline]{enumitem}
\usepackage[compact]{titlesec}

\setsansfont[BoldFont={Source Sans Pro Bold},
              Numbers={OldStyle}]{Source Sans Pro}
\setmainfont[BoldFont={Source Serif Pro Semibold},
              Numbers={OldStyle}]{Source Serif Pro}
\setmonofont{Source Code Pro}
\definecolor{linkColor}{RGB}{6,125,233}

\usepackage[dvipsnames]{xcolor}

\newcommand{\contact}[1]{{\color{red}#1}}

\usepackage[citestyle=numeric,backend=biber]{biblatex}
\bibliography{references}
\pagenumbering{gobble}

\titlespacing*{\section}{0em}{0.5em}{-0.5em}

\begin{document}
\renewcommand{\headrulewidth}{0pt}

\newenvironment{headerfont}{\color{gray}\fontfamily{ptm}\selectfont}{}
\pagestyle{fancy}{
  \lhead{\begin{headerfont}Stanford University\end{headerfont}}
  \chead{\begin{headerfont}\textbf{\underline{Waiver of Documentation}}\end{headerfont}}
  \rhead{\textbf{Approval Date}: \underline{Tentative}}
}

\section*{{\LARGE Managing crowd and gig workers' histories and qualifications}}

\section*{What's this study about?}
We're trying to learn about
whether independently--run platforms for tracking workers' histories
and qualifications
can better inform whether workers are likely
to be qualified and suited for complex gig work.
Some of this research will involve
the management of workers \textbf{by} other workers.

In this study, we'll be asking workers managed by this system about
their experiences using the platform (``vitae''), as well as
interviewing people who manage these qualification groups.
% the relationships online performers have with each other and
% with the
% organizations that run
% the platforms that they stream and
% upload to.
% Some of this touches on things like
% dealing with inter--personal issues and
% how the community deals with major events
% as well as questions about what it's like to
% make your income from online public performances.

\section*{What are my choices?}
You can choose not to participate in the study at all, or
you can respond to some questions but not others.
You can later ask that I not use some or
any of the information from any previous conversations
(until that data,
like an anonymized quotation, 
is published in a form that I can't change after the fact).
If you tell me you want to withdraw or
simply not participate in the first place,
I'll delete the data you gave
(if any) and that'll be the end of it.

\section*{Will anyone know I'm in the study?}
Only if you want ---
the only people in our research group who will know individual participants' names
will be people directly involved in the study.
The only reason other people will know you participated will be if
you tell them, or you tell me
to mention you when I email them.

\section*{What else should I know?}
If you're interested in participating, you should know
\begin{compactitem}
  \item Participation is voluntary; if you wish not to participate, it's perfectly all right.
  \item Participants will not be compensated for their interviewing time.
  \item If at any point you wish to withdraw from the study,
        we will accommodate such requests as quickly as reasonably possible.
  % \item Since it would be too difficult to get a signature from participants,
  %       (to say nothing of verifying the signature),
  %       we're using a ``\textbf{waiver of documentation}'',
  %       so this is purely informational.
  % \item The first conversation will probably take 30--45 minutes.
  %       Later conversations (if you're okay with it) would likely be shorter
  %       (mostly to clarify things or ask follow--up questions).
  % \item I can't compensate you (or anyone, really) for their time.
  % \item I plan on recording the audio of our conversation
  %       so I can transcribe what you said
  %       (so I don't misremember or misquote anything).
  %       Once I have a transcript, I'll delete the audio.
\end{compactitem}

\section*{Who can I talk to about the study?}
If you have any questions or concerns with the study,
you can talk to the researchers directly, or
you can talk to Stanford's Institutional Review Board (IRB),
who oversee our research and make sure participants are treated ethically.

\textbf{Talk to the researchers}:
If you have any questions or problems related to the study,
feel free to contact me, \contact{Ali Alkhatib},
by phone at \contact{408--348--2720} or
by email (\contact{\url{ali.alkhatib@cs.stanford.edu}}).

\textbf{Talk to the IRB}:
If you're not satisfied with how this study is being conducted, or
if you have any concerns, complaints, or
questions about the research or
your rights as a participant, please feel free to contact the
(IRB) to speak to someone independent of the research team at
\contact{(650)--723--2480} or
toll free at \contact{1--866--680--2906}.
You can also email \url{irb2-manager@lists.stanford.edu} or
write to:

\begin{center}\contact{Stanford IRB, Stanford University,\\
         3000 El Camino Real, Five Palo Alto Square, 4th Floor,\\
         Palo Alto, CA 94306}\end{center}

\end{document}